### 2024-08-16 article.tex
Share the draft

\documentclass[12pt,prb,aps]{revtex4-2}
\usepackage{amsmath,amssymb,physics}
\usepackage{hyperref} % Clickable links in the document
\usepackage{graphicx} % For figures with images
\usepackage{natbib} % bibliography
\setcitestyle{super, sort&compress, comma} % chemistry style
\usepackage{hyperref} % Clickable links in the document and their colors
\hypersetup{
    colorlinks=true, % false: boxed links; true: colored links
    linkcolor=cyan, % color of internal links
    citecolor=magenta, % color of links to bibliography
    filecolor=magenta, % color of file links
    urlcolor=cyan, % color of external links
    runcolor=cyan
}

\begin{document}

\title{Vibronic effects in the photoelectron spectrum of ozone}

% \author{Pawe{\l} W{\'o}jcik}
% \email[Electronic address: ]{pawel.wojcik@usc.edu}
% \affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}
% \affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE E{\"o}tv{\"o}s Lor{\'a}nd University, H-1518 Budapest 112, Hungary}
% \author{Anna I. Krylov}
% \email[Electronic address: ]{krylov@usc.edu}
% \affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}
% \author{John F. Stanton}
% \affiliation{Quantum Theory Project, Department of Chemistry, University of Florida, Gainesville, Florida 32611, USA}
% \affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE E{\"o}tv{\"o}s Lor{\'a}nd University, H-1518 Budapest 112, Hungary}

\date{\today}

\begin{abstract}Abstract\end{abstract}

% TODO: Add that in this work we use the Mulliken's notation for state
% symmetry

\maketitle

\section{Introduction}

Ozone is a C$_{2v}$ molecule with three normal modes: symmetric stretch,
symmetric bend, and asymmetric stretch. The asymmetric stretch is of a b$_2$
symmetry. The two lowest electronic states of the ozone cation are close in
energy. The lower state is fully symmetric while the higher one is of the
B$_2$ symmetry, the same symmetry as the asymmetric stretch. The close
energetic separation and the matching symmetry of the asymmetric stretch
results in significant vibronic effects appearing in the ozone photoelectron
spectrum. Figure~\ref{fig:ozone_intro} presents a graphical summary of this
information.

\begin{figure}
    \includegraphics[width = 16 cm]{./figures/ozone_intro}
    \caption{ 
        Two lowest states of the ozone cation and normal modes of ozone.
    }
    \label{fig:ozone_intro}
\end{figure}

\section{Methods}

We simulate the vibronic states of the ozone cation using the model
Hamiltonian of K{\"o}ppel, Domcke, and Cederbaum (KDC
Hamiltonian).\cite{Cederbaum:LVC:84,KDC:81,Koppel:CIbookCh7:04} This is a
multi-state and multi-mode Hamiltonian defined in the basis of diabatic
states.  We consider the ozone cation in the basis of two quasi-diabatic
states coupled by one mode (mode number 3). The model also includes two
symmetric modes (modes number 1 and 2)
\begin{subequations}
    \begin{equation}
        H = H _0 \mathbf{1}
        +
        \begin{pmatrix}
            V ^{(1)}  & \lambda _3 Q _3\\
            \lambda _3 Q _3 & V ^{(2)}
        \end{pmatrix}
    \end{equation}
    \begin{equation}
        H _0 = 
        \frac{1}{2} \left(\sum _{i = 1}^3 
        - \omega _i \frac{\partial ^2}{\partial Q _i ^2 }\right)
        + \frac{1}{2}\omega _3 Q _3 ^2
    \end{equation}
    \begin{equation}
        V ^{(\alpha)} = 
        E ^{(\alpha)} 
        + \sum _{i,j,k,l \in \{1,2\}} 
            \kappa ^{(\alpha)} _i Q _i 
            + \kappa ^{(\alpha)} _{ij} Q _i Q _j 
            + \kappa ^{(\alpha)} _{ijk} Q _i Q _j Q _k 
            + \kappa ^{(\alpha)} _{ijkl} Q_i Q _j Q _k Q _l.
    \end{equation}
    \label{eq:KDC}
\end{subequations}
The Hamiltonian parameters are expanded around the equilibrium geometry of
ozone. $E ^{(\alpha)}$ are the vertical ionization energies calculated at that
geometry. $Q_i$ are the dimensionless normal coordinates of ozone. $\kappa$ are
the coefficients of expansion of the potential along the fully symmetric
coordinates. $\lambda$ are the linear diabatic couplings.  $\omega$ are the
harmonic frequencies of ozone.

We find the parameters that enter the KDC Hamiltonian using \emph{ab inito}
coupled-cluster (CC) methods and its equation-of-motion EOM-CC
extension.~\cite{Bartlett:CC_review:07, Krylov:EOMRev:07, Bartlett:Book:09,
Christiansen:EOMRev:11, Bartlet:EOMRev:12, Krylov:OSRev} We use the CC
truncated to singles and doubles (CCSD), singles, doubles and triples (CCSDT)
as well as singles, doubles, triples and quadruples
(CCSDTQ).\cite{Matthews:ncc:2015}  We use the EOM-CC for ionization potential
(EOM-IP).~\cite{StantonGauss:EOMIP:99} We use the Ichino, Gauss and Stanton
definition of quasi-diabatic states based on the EOM-CC method
(EOM-CC-QD)~\cite{Stanton:EOMIPdeg:09}. In all CC and EOM-CC calculations we
leave all electrons correlated, in other words, we do not use the
frozen-core approximation.

Using CCSDT/ANO1 we optimize the geometry of ozone and compute its harmonic
frequencies and normal coordinates.~\cite{Almlof:ANO,Almlof:ANO:1988} At the
same geometry we compute the linear diabatic coupling $\lambda$ using
EOM-IP-CCSD-QD/ANO1.  We use EOM-IP-CCSDT/ANO1 on a grid to find the
expansion coefficients $\kappa$.

The vertical ionization energies are calculated using a composite method. The
base value is the complete basis set (CBS) extrapolation of the
EOM-IP-CCSDT/cc-pCVnZ energies with n = 5, 6.~\cite{Woon:95:CCBS} These values
are augmented with two corrections: the $\Delta$Q correction in the cc-pwCVTZ
basis set and the relativistic correction calculated using
EOM-IP-CCSD/cc-pwCVTZ.~\cite{Dunning:02:p(w)CVnZ} We introduce an error
estimates to the reported vertical energies. For the error estimate of the
extrapolated CBS value we use half of the absolute value of the difference
between the best \emph{ab initio} value and the extrapolated value. For the
remaining corrections we use as the error estimate half of the absolute value
of the correction.

To compare the simulated photoelectron spectrum to the experimental one, we
use the oscillator strengths ratio A$_1$:B$_2$ of $1$:$1.35$.\cite{KDC:O3:92}
Additionally, the stick spectrum is broadened with the Lorenzian envelopes
normalized to the peaks' intensities
\begin{equation}
    f _{envelope}(x, x _i, I _i) = 
    \frac{I_i}{(x-x _i)^2 + (\gamma/ 2) ^2}.
    \label{eq:lorentzian}
\end{equation}
$x _i$ is the position of the spectral peak, $I _i$ is its intensity and
$\gamma$ is the peak's width.

We use the \textsc{xsim} program of Dr. Stanton to simulate the spectrum using
the basis of 50 harmonic states in each mode and 6000 iterations of the
Lanczos procedure. All remaining \emph{ab initio} calculations are completed
using \textsc{CFOUR}.~\cite{cfour, cfour:2020}

\section{Results}

The optimized geometry of ozone gives the bond lengths of $1.270$~\AA{} and
the bond angle of $116.9^\circ$. The two symmetric normal modes have
frequencies $1169$~cm$^{-1}$ and $724$~cm$^{-1}$, while the asymmetric stretch
has frequency equal to $1129$~cm$^{-1}$. The computed value of the linear
diabatic coupling constant $\lambda$ is $1394$~cm$^{-1}$. The vertical
ionization energy for the first excited state, $E^{(1)}$, computed using the
composite method described above is equal to $12.827$~eV. Our error estimate
for this value is $30$~meV, see Table~\ref{tab:vertical_ionization_energy} for
details.  We note that the convergence of the vertical ionization gap between
the two states is much faster and we estimate this value as equal to
$123\pm8$~meV, see Table~\ref{tab:vertical_gap} for details.

\begin{table}
    \caption{
        Vertical ionization energies with the error estimates, eV.
    }
    \label{tab:vertical_ionization_energy}
    \begin{center}
        \begin{tabular}[c]{|l|rr|r|}
            \hline
            Contribution  & $^2$A$_1$ & $^2$B$_2$ & Error estimate \\ \hline
            CBS           & 12.872    & 12.981    & 0.02 \\
            $\Delta$Q     & -0.037    & -0.021    & 0.02 \\
            Relativistic  & -0.008    & -0.010    & 0.005 \\ \hline
            Sum           & 12.827    & 12.950    & 0.03 \\ \hline
        \end{tabular}
    \end{center}
\end{table}

\begin{table}
    \caption{
        The energy gap between vertical ionization energies of the $^2$A$_1$
        and $^2$B$_2$ states with error estimates, meV. See caption of
        Table~\ref{tab:vertical_ionization_energy}.
    }
    \label{tab:vertical_gap}
    \begin{center}
        \begin{tabular}[c]{|l|r|r|}
            \hline
            Contribution             &  Gap    & Error estimate \\ \hline
            EOM-CCSDT/CBS            &  108.8  & 0.9 \\
            $\Delta$Q/pwCVTZ         &   16.5  & 8 \\
            Relativistic/CCSD/pwCVTZ &   -2.2  & 1.1 \\ \hline
            Final value, meV         &  123    & 8 \\
            Final value, cm$^{-1}$   &  990    & 65 \\ \hline
        \end{tabular}
    \end{center}
\end{table}

\begin{figure}
\includegraphics[width = 8 cm]{figures/sim_vs_Dyke}
\caption{
    Comparison of the experimental (solid black line) and the simulated
    spectrum. In the simulated spectra the plot is generated with peak width
    $\gamma = 30$~meV. The simulated spectrum is shifted towards higher
    energies by $21$~meV.
}
\label{fig:sim_vs_dyke}
\end{figure}

Figure~\ref{fig:sim_vs_dyke} compares the simulated spectrum to the
experimental one taken from reference~\cite{dyke:O3:74}. A comparison of the
simulated and experimental spectra is presented on
Figure~\ref{fig:sim_vs_dyke} and shows a good match. Peak A is known to be a
hot band.~\cite{KDC:O3:92} The simulation reproduces well the consecutive
increase in the intensities of peaks B, C, D, and E. The spacing between these
peaks is also well reproduced. Drop in the intensity of peak F is also
captured by the simulation. Starting from peak G the simulation shows
discrepancy with the experiment. A sudden drop in the intensity past the peak
H is not observed in the simulation. We discuss a likely source of this
mismatch later.

\begin{figure}
    \includegraphics[width=10 cm]{figures/spectrum_overline}
    \caption{
        Simulated photoelectron spectrum of ozone. Bottom axis shows energy
        scale in eV. Top axis shows energy offset from the origin in
        cm$^{-1}$. Stick spectrum shows positions and intensities of all
        simulated states. Blue color corresponds to the oscillator strength
        originating from the $^2$A$_1$ basis state while the red ones
        correspond to $^2$B$_2$. Gray vertical lines with captions on top
        indicate positions of features as measured by the PFI-ZEKE
        experiment.~\cite{Willitsch:O3ZEKE:2005} 
        $D_0$ marks the dissociation threshold of O$_3^+$.
        Gray dotted line marks the energy of the minimum of the conical
        intersection (CI).
    }
    \label{fig:ozone_overlay}
\end{figure}

Our simulation allows for an additional element of analysis of the simulated
spectrum. Figure~\ref{fig:ozone_overlay} presents the same simulated spectrum
as in Figure~\ref{fig:sim_vs_dyke} with additional information on the
decomposition of the spectrum.  All lines that contribute to the spectrum are
marked individually and in two different colors. The spectrum shows that peaks
A and B are almost purely of the $^2$A$_1$ character. Starting from peak C
each the contributions from two states are equally important, it is also clear
that each peak of the low-resolution spectrum has in fact contributions from
many vibronic peaks. At the energy of about $2000$~cm$^{-1}$ above the origin
the density of vibronic peaks increases significantly. This value can be
compared to the minimum of the conical intersection, which our simulation
locates at $3174$~cm$^{-1}$ above the origin ($12.92$~eV). Our simulation is
incapable of accounting for dissociation of the molecule, therefore the we
expect a discrepancy with the experiment as the energy gets closer to the
dissociation threshold located at $4840$~cm$^{-1}$ above the
origin.~\cite{Willitsch:O3ZEKE:2005}

\begin{figure}
    \includegraphics[width=10cm]{figures/spectrum_assigned.pdf}
    \caption{
        Simulated photoelectron spectrum of ozone, but with no account given
        to the vibronic coupling effect.
    }
    \label{fig:no_coupling}
\end{figure}

We would like to assign the vibronic peaks from our simulation. To this end,
we run the simulation once again, this time, however we set the linear
diabatic constant to zero, i.e., $\lambda = 0$. With that change we can
reproduce the spectrum (at an equivalent level of theory) but in an artificial
case where there is no vibronic coupling. Figure~\ref{fig:no_coupling}
presents this spectrum. The non-coupled spectrum is easy to assign using the
labels that mark the symmetry of the electronic state, A$_1$ or B$_2$, and the
vibrational state ($\nu _1 \nu_2 \nu_3$), where $\nu _i$ is the number of
quanta in mode $i$ with $i=1$ for the symmetric stretch, $i=2$ for the
symmetric bend and $i=3$ for the asymmetric stretch. The assigned spectrum
shows progressions in the symmetric bend. There is one such progression in
each electronic states.  Additionally for each state, there is also another
progression in the symmetric bend with one excitation in the symmetric
stretch.

\begin{figure}
    \includegraphics[width=16 cm]{./figures/sim_vs_zeke}
    \caption{
        Comparison of the simulated spectrum with the PFI-ZEKE
        experiment.\cite{Willitsch:O3ZEKE:2005} The simulated spectrum was
        shifted by $21$~meV$ = 170$~cm$^{-1}$ towards higher energies.
    }
    \label{fig:sim_vs_zeke}
\end{figure}

We decompose the vibronic states of our main simulation in the basis of the
artificial, uncoupled states discussed in the previous paragraph.
Figure~\ref{fig:sim_vs_zeke} shows the assigned spectrum and
Table~\ref{tab:peak_assignment} lists the decomposition of all peaks in the
region of up to $3000$~cm$^{-1}$ with intensities larger than $10^{-3}$. We
compare the assigned spectrum against the high-resolution
pulsed-field-ionization zero-kinetic-energy (PFI-ZEKE) spectrum from
2005.\cite{Willitsch:O3ZEKE:2005} The PFI-ZEKE spectrum is the best source of
information on the location of the peaks, especially the origin, against which
align our simulation. The origin of the simulated spectrum is located
$170$~cm$^{-1}$ lower than the experimental origin which was experimentally
observed at $101,020.5$~cm$^{-1}$.\cite{Willitsch:O3ZEKE:2005} This is within
our error estimate for the vertical ionization energy
($30$~meV$=240$~cm$^{-1}$).  Once the origin locations are aligned the
comparison shows a good match between the simulation and the experiment.
Especially the states with an oscillator strength originating from the A$_1$
state are well aligned with the experimental features.

States close to the origin of the band are similar to the non-coupled states.
The origin peak is of a clear A$_1(0,0,0)$ character, the first two
excitations in the symmetric bend, A$_1(0,1,0)$ and A$_1(0,2,0)$ are also very
similar to the non-coupled states, while the higher excitations in this
progression are showing large mixing. The same progression with one
vibrational quantum in the symmetric stretch is more interesting. The first of
its peaks is very well aligned with an experimentally visible feature, which
was previously assigned as the origin of the second state. The second peak in
this series, the first combination state A$_1(1,1,0)$, has very high
similarity to its uncoupled version. It is a good candidate to assign the
experimental, unassigned feature above $102,600$~cm$^{-1}$.

States with an oscillator strength from the B$_2$ state are on the other hand
not aligning well with the experiment. First such peak, close to the $12.64$~eV
mark on the Figure~\ref{fig:sim_vs_zeke}, is of a vibronic character and we
assign it as a A$_1$(0,0,1) state which gains intensity thanks to the coupling
to the B$_2$ state. This vibronic peak lies in the part of the spectrum marked
as pollution due to water. The origin of the B$_2$ state is very similar to
the uncoupled B$_2(0,0,0)$ state, but it is located in an empty area of the
experimental spectrum. The next peak, slightly above the $12.76$~eV mark on
Figure~\ref{fig:sim_vs_zeke}, corresponds to the one excitation of the symmetric
bend in the B$_2$ state, but as the label on the figure shows, it is only
about $55$\% similar to the uncoupled state.

\begin{figure}
    \includegraphics[width=16 cm]{figures/sim_vs_TarroniCarter}
    \caption{
        Comparison of the simulated spectrum to an earlier accurate simulation
        of Tarroni and Carter (assigned lines).~\cite{tarroni:O3:2011}
    }
    \label{fig:sim_vs_tarronicarter}
\end{figure}

Finally, we compare our simulation to an earlier accurate simulation, the work
of Tarroni and Carter from 2011.~\cite{tarroni:O3:2011} Both spectra show a
good match to the experiment and agree in the assignment of the first
progression in the bending mode. The previous simulation, however, shows
smaller spacing between other lines, showing much higher congestion of the
spectrum. Additionally the origin of the second state falls at lower energy
aligning well with the peak that was assigned in the PFI-ZEKE also as the
origin of the second state.  While an additional simulation of the line
intensities would allow for the most complete comparison to the experimental
spectrum, the strong alignment of the simulated peaks of a leading
A$_1$ character with the experimental features leads us to believe that our
simulation offers the best assignment of the spectrum to date.

\begin{table}
    \renewcommand{\arraystretch}{0.8}% Tighter
    \caption{Assignment and decomposition of the eigenvectors of the ozone
    cation.}
    \label{tab:peak_assignment}
    \begin{tabular}{|l|l|l|}
        \hline
        Peak position (cm$^{-1}$) & Assignment & Eigenvector \\
        \hline
$     0$ &  $A_1(000)$ & $ -0.99~A_1(000)$ \\
$   618$ &  $A_1(010)$ & $ +0.99~A_1(010)$ \\
$   915$ &  $A_1(001)$ & \\
$  1076$ &  $A_1(100)$ & $ -0.99~A_1(100)$ \\
$  1222$ &  $A_1(020)$ & $ -0.97~A_1(020)$ \\
$  1368$ &  $B_2(000)$ & $ +0.97~B_2(000)$ \\
$  1498$ &             & $ -0.10~B_2(000)$ \\
$  1684$ &  $A_1(110)$ & $ -0.96~A_1(110)$ \\
$  1796$ &  $A_1(030)$ & $ -0.87~A_1(030) -0.16~A_1(110) -0.13~A_1(020) +0.13~A_1(040)$ \\
$  1848$ &             & $ +0.31~A_1(030)$ \\
$  1921$ &  $B_2(010)$ & $ -0.74~B_2(010) +0.18~B_2(020) +0.11~B_2(000)$ \\
$  1977$ &             & $ +0.24~B_2(010)$ \\
$  2085$ &             & $ -0.52~B_2(010)$ \\
$  2132$ &             & $ +0.25~A_1(030) +0.24~A_1(040) +0.13~A_1(120) -0.13~A_1(050)$ \\
$  2275$ &  $A_1(120)$ & $ +0.86~A_1(120) +0.12~A_1(040)$ \\
$  2363$ &             & $ -0.62~A_1(040) +0.40~A_1(120) -0.17~A_1(030) +0.11~A_1(050)$ \\
$  2439$ &             & $ +0.47~A_1(040) +0.12~A_1(120)$ \\
$  2454$ &             & $ +0.53~B_2(020) +0.24~B_2(010) -0.17~B_2(030)$ \\
$  2576$ &             & $ +0.27~B_2(020)$ \\
$  2736$ &             & $ +0.60~B_2(020)$ \\
$  2886$ &  $A_1(130)$ & $ -0.81~A_1(130) +0.12~A_1(050) -0.10~A_1(120)$ \\
$  2974$ &             & $ -0.24~B_2(110) -0.16~B_2(020)$ \\
$  3047$ &  $A_1(050)$ & $ +0.76~A_1(050) +0.13~A_1(130)$ \\
        \hline
    \end{tabular}
\end{table}

\section{Summary and Conclusions}

We have simulated the vibronic effects in the photoelectron spectrum of the
ozone.  We have used the KDC Hamiltonian\cite{Cederbaum:LVC:84, KDC:81,
Koppel:CIbookCh7:04} with the \emph{ab initio} parametrisation of the
quasi-diabatic states of Ichino, Gauss and Stanton.~\cite{Stanton:EOMIPdeg:09}
The results of our simulation match well the photoelectron spectrum from
1970s~\cite{dyke:O3:74} and the PFI-ZEKE spectrum from
2005.~\cite{Willitsch:O3ZEKE:2005} We present analysis of the spectra and
highlight the role of the vibronic coupling effects. Our simulation offers a
state-of-the-art insight into the spectrum and allows us to assign the
simulated peaks. Our assignment agrees well with the previously assigned
progression in the symmetric bend but we reassign some of the other features
of the spectrum. Simulated peaks that gain intensity from the B$_2$ state are
absent in the PFI-ZEKE spectrum which offers an interesting avenue for further
investigations. Modeling of the intensity of the PFI-ZEKE peaks is not
unlikely to bring insight into this problem.  Additionally our simulation does
not model dissociation of the ozone cation, which will play a role in the high
energy part of the spectrum.

\section{Acknowledgments}

This work was funded by the NSF Center for Chemical Innovation Phase I (grant
no. CHE-2221453) and U.S. Department of Energy, Basic Energy Sciences (grant
no. DE-FG02-05ER15629).

\clearpage

\bibliographystyle{prf}
\bibliography{allrefs}

\end{document}

% vim: tw=78 spell:

### 2024-08-17 article.tex
Merge introduction

@@
-\documentclass[12pt,prb,aps]{revtex4-2}
+\documentclass[12pt,pra,aps,superscriptaddress]{revtex4-2}
 \usepackage{amsmath,amssymb,physics}
@@
 
-\title{Vibronic effects in the photoelectron spectrum of ozone}
+\title{Vibronic Coupling Effects in the Photoelectron Spectrum of Ozone: A
+Coupled-Cluster Approach}
 
-% \author{Pawe{\l} W{\'o}jcik}
-% \email[Electronic address: ]{pawel.wojcik@usc.edu}
-% \affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}
-% \affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE E{\"o}tv{\"o}s Lor{\'a}nd University, H-1518 Budapest 112, Hungary}
-% \author{Anna I. Krylov}
-% \email[Electronic address: ]{krylov@usc.edu}
-% \affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}
-% \author{John F. Stanton}
-% \affiliation{Quantum Theory Project, Department of Chemistry, University of Florida, Gainesville, Florida 32611, USA}
-% \affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE E{\"o}tv{\"o}s Lor{\'a}nd University, H-1518 Budapest 112, Hungary}
+\author{Pawe{\l} W{\'o}jcik}
+\affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}
+\affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE
+E{\"o}tv{\"o}s Lor{\'a}nd University, P{\'a}zm{\'a}ny P{\'e}ter stny. 1/A,
+Budapest, Hungary}
+
+\author{Anna I. Krylov}
+\affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}
+
+\author{Hannah Reisler}
+\affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}
+
+\author{P{\'e}ter G. Szalay}
+\affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE
+E{\"o}tv{\"o}s Lor{\'a}nd University, P{\'a}zm{\'a}ny P{\'e}ter stny. 1/A,
+Budapest, Hungary}
+
+\author{John F. Stanton}
+\email[Electronic address: ]{jfstanton@gmail.com}
+\affiliation{Quantum Theory Project, Departments of Chemistry and Physics, University of Florida, Gainesville, FL, USA 32611}
+\affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE
+E{\"o}tv{\"o}s Lor{\'a}nd University, P{\'a}zm{\'a}ny P{\'e}ter stny. 1/A,
+Budapest, Hungary}
 
@@
 
-\begin{abstract}Abstract\end{abstract}
-
-% TODO: Add that in this work we use the Mulliken's notation for state
-% symmetry
+\begin{abstract}
+One of the most important areas of application of equation-of-motion coupled
+cluster (EOM-CC) theory is to the prediction, simulation and analysis of
+various types of electronic spectra.   In this work, the EOM-CC method known
+as EOMIP-CC is applied to the closely lying and coupled pair of states of the
+ozone cation -- ${\tilde X}^2A_1$ and ${\tilde A}^2B_2$ -- using sophisticated
+approaches that extend to the full singles, doubles, triples and quadruples
+model (EOMIP-CCSDTQ).   Combined with a venerable and powerful method for
+calculating vibronic spectra from the Hamiltonian produced by EOMIP-CC
+calculations, this research provides a spectrum that is in good agreement with
+the photoelectron spectrum of ozone.   An important result here is that the
+calculations suggest that the adiabatic gap separating these two electronic
+states is somewhat smaller than currently thought; an assignment of the
+simulated spectrum together with the more precise band positions of the
+experimental measurements suggests that this energy gap is 1366$\pm$65
+cm$^{-1}$.
+\end{abstract}
 
@@
 
-Ozone is a C$_{2v}$ molecule with three normal modes: symmetric stretch,
-symmetric bend, and asymmetric stretch. The asymmetric stretch is of a b$_2$
-symmetry. The two lowest electronic states of the ozone cation are close in
-energy. The lower state is fully symmetric while the higher one is of the
-B$_2$ symmetry, the same symmetry as the asymmetric stretch. The close
-energetic separation and the matching symmetry of the asymmetric stretch
-results in significant vibronic effects appearing in the ozone photoelectron
-spectrum. Figure~\ref{fig:ozone_intro} presents a graphical summary of this
-information.
+Amongst the brotherhood of triatomic molecules, it cannot be argued that water
+(H$_2$O) is the most important, the most highly studied, and the most well
+understood.  Beyond H$_2$O, there are many triatomic molecules that have an
+environmental, technological or biological importance, have been subjected to
+many studied and are understood to various levels of detail.  Perhaps the most
+interesting such case is ozone (O$_3$), which has a vast number of important
+properties, a very rich history of study \cite{chappuis}, and – unlike the
+relatively simple water molecule – a profound quantum-mechanical complexity
+\cite{Babikov:anomalousOzone:2003}.  In the latter context, while we think of
+(and an NMR experiment would reveal) as two distinct kinds of oxygen Atoms in
+ozone, the full molecular Hamiltonian does not distinguish between the two
+types, with the three equivalent structures separated by a barrier that lies
+tantalizingly close to the O$_3$ $\rightarrow$ O$_2$ + O dissociation
+threshold (102.4 kJ/mol) \cite{Ruscis:ATcT:2022}. In reality, the energy
+levels of ozone all have a near triple ($e+a$) degeneracy, albeit with a
+tunneling splitting so small that it can be ignored, along with a
+semi-infinite lifetime (despite opposing opinion
+\cite{Boggs:BerryOzone:2006}).   More than a half century ago, this intriguing
+aspect of the ozone molecule was first discussed by Berry
+\cite{Berry:Ozone:1960}.
+
+Beyond the structural aspects of ozone, other mysteries center around this
+curious molecule.  For example, the distribution of the eighteen distinct
+$^{16}$O/$^{17}$O/$^{18}$O isotopologues in the Earth’s atmosphere differs
+from what is expected based on the natural isotopic abundance, a puzzle that
+has been open for more than three decades
+\cite{Mauersberger:OzoneMystery:1990}.
+
+Amongst quantum chemists, ozone has a notorious history, with its strong
+biradical character causing significant difficulties in calculations of its
+ostensibly simple ground state molecular properties.  An early 1989 study
+\cite{Stanton:Ozone:1989, Stanton:Ozone:1989b} by the Bartlett group and
+collaborators found that the CCSD+T(CCSD) method predicted that the molecular
+equilibrium structure of ozone would have $C_s$ symmetry (that is, the
+asymmetric stretching harmonic frequency predicted by this method was
+imaginary), a finding that led to a search for better treatment of
+non-iterative triple excitations, ultimately leading to the well-known CCSD(T)
+treatment \cite{Raghavachari:89, Urban:ccsd(t):1985, GenrefCCSD(T):93}. While
+CCSD(T) and higher-level coupled-cluster methods available today
+\cite{kucharski:ccsdtq:1992, Kallay:CCHigh, Matthews:ncc:2015} do a good job
+in predicting the molecular vibrational potential, an elaborate multireference
+configuration interaction study by Dawes {\it et al.} has done an excellent
+job on the entire ground-state surface out to the dissociation limit
+\cite{Dawes:ozone:2013}. As such, the quantum chemical understanding and
+fidelity for the ground electronic state of O$_3$ is now at a mature level.
+
+Qualitatively, the challenge posed to electronic structure by ozone ultimately
+arises from its closely spaced highest-occupied ($b_2$) and lowest-unoccupied
+($a_1$) molecular orbitals (HOMO and LUMO, respectively).  The two electron
+configurations $[\cdots]b_2^2 a_1^0$ and $[\cdots]b_2^0 a_1^2$ mix strongly,
+posing the aforementioned challenges with (especially single-reference)
+quantum chemical methods.   Another consequence of the identity and energetic
+proximity of these two orbitals is that the ozone cation (which is
+isoelectronic to the NO$_2$ radical) has closely lying $^2A_1$ and $^2B_2$
+electronic states.   Like the associated states in NO$_2$, both of these
+states are plagued by orbital symmetry-breaking problems
+\cite{Davidson:SymmBreak:76}, a problem that greatly complicates
+quantum-chemical efforts to them.   One of the many accomplishments of the
+Bartlett group has been the integral role played by them in the development of
+equation-of-motion coupled-cluster theory \cite{Stanton:93:EOMCC,
+Nooijen:EOMEA:95, Bartlett:CC_review:07} (EOM-CC, also known as linear
+response coupled-cluster theory \cite{Koch:90:LinResp}). These methods provide
+a very efficient and simple way to study certain classes of what are often
+termed ``multireference problems" \cite{Krylov:EOMRev:07}, and are ideally
+suited to studying many reactive intermediates, radicals, biradicals and
+electronically excited states.   From a somewhat wider viewpoint, the
+existence of closely spaced electronic states always carries the potential for
+(possibly strong) vibronic coupling, a phenomenon that can play an important
+role in molecular dynamics and spectroscopy.  Indeed, one of the great
+successes of EOM-CC methods has been in their ability -- if and only if
+combined with vibronic coupling models -- to enable high-quality simulations
+of complicated electronic spectra.   Such work provides important insight into
+the nature of vibronic coupling in molecular systems, as has been exemplified
+by various application studies (for example, see Refs.~\cite{Stanton:NO3:07}
+and~\cite{Koppel:02}).
+
+Our contribution to this issue paying homage to the career and accomplishments
+of R.J. Bartlett consists of an application of a vibronic coupling model
+parametrized by EOM-CC calculations to the photoelectron spectrum of ozone.
+We trust that this combination of methodology applied to a molecule that has
+been extensively studied by the Bartlett group is an appropriate contribution
+to this issue.
 
@@
 \section{Methods}
+
+Ozone is a $C_{2v}$ molecule (we are following the Mulliken's
+convention~\cite{Mulliken:55:symnot} by placing the molecule in the $yz$ plane
+and aligning the molecular symmetry axis with the $z$ direction) with three
+normal modes: symmetric stretch, symmetric bend, and asymmetric stretch.  The
+asymmetric stretch is of a $b_2$ symmetry. The two lowest electronic states of
+the ozone cation are close in energy. The lower state is fully symmetric while
+the higher one is of the $B_2$ symmetry, the same symmetry as the asymmetric
+stretch. The close energetic separation and the matching symmetry of the
+asymmetric stretch results in significant vibronic effects appearing in the
+ozone photoelectron spectrum.  Figure~\ref{fig:ozone_intro} presents a
+graphical summary of this information.
 
@@
 
-This work was funded by the NSF Center for Chemical Innovation Phase I (grant
-no. CHE-2221453) and U.S. Department of Energy, Basic Energy Sciences (grant
-no. DE-FG02-05ER15629).
+This project was initiated when three of the authors (P.W, P.G.S. and J.F.S)
+were in Budapest, where J.F.S. was serving as a recipient of the John von
+Neumann Award in STEM bestowed by the Fulbright Foundation. Additional
+research presented here benefited from the NSF Center for Chemical Innovation
+Phase I (grant no. CHE-2221453) and U.S. Department of Energy, Basic Energy
+Sciences (grant no. DE-FG02-05ER15629).  All the authors of this research wish
+Prof. Bartlett a happy ninth decade of life, and hope that he continues to
+stimulate others in the field with his creative insights.
 

### 2024-08-17 article.tex
Split Results into Results and Discussion

@@
-\documentclass[12pt,pra,aps,superscriptaddress]{revtex4-2}
+\documentclass[12pt,a4paper,prb,aps,superscriptaddress]{revtex4}
 \usepackage{amsmath,amssymb,physics}
-\usepackage{hyperref} % Clickable links in the document
 \usepackage{graphicx} % For figures with images
@@
 \author{John F. Stanton}
-\email[Electronic address: ]{jfstanton@gmail.com}
+\email[Electronic address: ]{jfstanton137@gmail.com}
 \affiliation{Quantum Theory Project, Departments of Chemistry and Physics, University of Florida, Gainesville, FL, USA 32611}
@@
 
-\section{Methods}
+\section{Theory}
 
@@
 Figure~\ref{fig:sim_vs_dyke} compares the simulated spectrum to the
-experimental one taken from reference~\cite{dyke:O3:74}. A comparison of the
-simulated and experimental spectra is presented on
-Figure~\ref{fig:sim_vs_dyke} and shows a good match. Peak A is known to be a
-hot band.~\cite{KDC:O3:92} The simulation reproduces well the consecutive
-increase in the intensities of peaks B, C, D, and E. The spacing between these
-peaks is also well reproduced. Drop in the intensity of peak F is also
-captured by the simulation. Starting from peak G the simulation shows
-discrepancy with the experiment. A sudden drop in the intensity past the peak
-H is not observed in the simulation. We discuss a likely source of this
-mismatch later.
+experimental one taken from reference~\cite{dyke:O3:74}. Our simulation allows
+for an additional element of analysis of the simulated spectrum.
+Figure~\ref{fig:ozone_overlay} presents a decomposition of the simulated
+spectrum from Figure~\ref{fig:sim_vs_dyke}. All lines that contribute to the
+spectrum are marked individually and are color-coded indicating which
+electronic state's transition intensity the peak draws from. The total
+envelope of the spectrum is also decomposed showing contributions from both
+states. Our simulation locates the minimum of the conical intersection at
+$3174$~cm$^{-1}$ above the origin ($12.92$~eV) that is marked on the figure as
+CI.
 
@@
         simulated states. Blue color corresponds to the oscillator strength
-        originating from the $^2$A$_1$ basis state while the red ones
-        correspond to $^2$B$_2$. Gray vertical lines with captions on top
+        originating from the $^2$A$_1$ basis state while the red one
+        corresponds to $^2$B$_2$. Gray vertical lines with captions on top
         indicate positions of features as measured by the PFI-ZEKE
-        experiment.~\cite{Willitsch:O3ZEKE:2005} 
-        $D_0$ marks the dissociation threshold of O$_3^+$.
-        Gray dotted line marks the energy of the minimum of the conical
-        intersection (CI).
+        experiment.~\cite{Willitsch:O3ZEKE:2005} The simulated spectrum was
+        shifted to match the PFI-ZEKE experimental origin. $D_0$ marks the
+        dissociation threshold of O$_3^+$. Gray dotted line marks the energy
+        of the minimum of the conical intersection (CI).
     }
@@
 \end{figure}
-
-Our simulation allows for an additional element of analysis of the simulated
-spectrum. Figure~\ref{fig:ozone_overlay} presents the same simulated spectrum
-as in Figure~\ref{fig:sim_vs_dyke} with additional information on the
-decomposition of the spectrum.  All lines that contribute to the spectrum are
-marked individually and in two different colors. The spectrum shows that peaks
-A and B are almost purely of the $^2$A$_1$ character. Starting from peak C
-each the contributions from two states are equally important, it is also clear
-that each peak of the low-resolution spectrum has in fact contributions from
-many vibronic peaks. At the energy of about $2000$~cm$^{-1}$ above the origin
-the density of vibronic peaks increases significantly. This value can be
-compared to the minimum of the conical intersection, which our simulation
-locates at $3174$~cm$^{-1}$ above the origin ($12.92$~eV). Our simulation is
-incapable of accounting for dissociation of the molecule, therefore the we
-expect a discrepancy with the experiment as the energy gets closer to the
-dissociation threshold located at $4840$~cm$^{-1}$ above the
-origin.~\cite{Willitsch:O3ZEKE:2005}
 
@@
 shows progressions in the symmetric bend. There is one such progression in
-each electronic states.  Additionally for each state, there is also another
+each electronic state.  Additionally for each state, there is also another
 progression in the symmetric bend with one excitation in the symmetric
@@
 our error estimate for the vertical ionization energy
-($30$~meV$=240$~cm$^{-1}$).  Once the origin locations are aligned the
-comparison shows a good match between the simulation and the experiment.
-Especially the states with an oscillator strength originating from the A$_1$
-state are well aligned with the experimental features.
-
-States close to the origin of the band are similar to the non-coupled states.
-The origin peak is of a clear A$_1(0,0,0)$ character, the first two
-excitations in the symmetric bend, A$_1(0,1,0)$ and A$_1(0,2,0)$ are also very
-similar to the non-coupled states, while the higher excitations in this
-progression are showing large mixing. The same progression with one
-vibrational quantum in the symmetric stretch is more interesting. The first of
-its peaks is very well aligned with an experimentally visible feature, which
-was previously assigned as the origin of the second state. The second peak in
-this series, the first combination state A$_1(1,1,0)$, has very high
-similarity to its uncoupled version. It is a good candidate to assign the
-experimental, unassigned feature above $102,600$~cm$^{-1}$.
-
-States with an oscillator strength from the B$_2$ state are on the other hand
-not aligning well with the experiment. First such peak, close to the $12.64$~eV
-mark on the Figure~\ref{fig:sim_vs_zeke}, is of a vibronic character and we
-assign it as a A$_1$(0,0,1) state which gains intensity thanks to the coupling
-to the B$_2$ state. This vibronic peak lies in the part of the spectrum marked
-as pollution due to water. The origin of the B$_2$ state is very similar to
-the uncoupled B$_2(0,0,0)$ state, but it is located in an empty area of the
-experimental spectrum. The next peak, slightly above the $12.76$~eV mark on
-Figure~\ref{fig:sim_vs_zeke}, corresponds to the one excitation of the symmetric
-bend in the B$_2$ state, but as the label on the figure shows, it is only
-about $55$\% similar to the uncoupled state.
+($30$~meV$=240$~cm$^{-1}$).
 
@@
 
-Finally, we compare our simulation to an earlier accurate simulation, the work
-of Tarroni and Carter from 2011.~\cite{tarroni:O3:2011} Both spectra show a
-good match to the experiment and agree in the assignment of the first
-progression in the bending mode. The previous simulation, however, shows
-smaller spacing between other lines, showing much higher congestion of the
-spectrum. Additionally the origin of the second state falls at lower energy
-aligning well with the peak that was assigned in the PFI-ZEKE also as the
-origin of the second state.  While an additional simulation of the line
-intensities would allow for the most complete comparison to the experimental
-spectrum, the strong alignment of the simulated peaks of a leading
-A$_1$ character with the experimental features leads us to believe that our
-simulation offers the best assignment of the spectrum to date.
+Lastly, we compare our simulation to an earlier accurate simulation, the work
+of Tarroni and Carter from 2011.~\cite{tarroni:O3:2011}
+Figure~\ref{fig:sim_vs_tarronicarter} presents a comparison of lines from both
+simulations. The comparison includes only the lines from the earlier
+simulation that were tabulated with an assignment.
 
@@
 
+\section{Discussion}
+
+We first discuss the comparison of our simulations to the lower-resolution
+experimental spectrum presented on Figure~\ref{fig:sim_vs_dyke}. The peak A is
+known to be a hot band.~\cite{KDC:O3:92} The simulation reproduces well the
+consecutive increase in the intensities of peaks B, C, D, and E. The spacing
+between these peaks is also well reproduced. Drop in the intensity at the peak
+F is captured by the simulation. Starting from the peak G, the simulation
+shows discrepancy with the experiment. A sudden drop in the intensity past the
+peak H is not observed in the simulation. We discuss a likely source of this
+mismatch later.
+
+The decomposition of the spectrum presented on Figure~\ref{fig:ozone_overlay}
+introduces additional insight. The spectrum shows that peaks B and C are
+almost purely of the $^2$A$_1$ character. Starting from the peak D, the
+contributions from two states are equally important. It is also clear that the
+following peaks arise as mixtures of many vibronic peaks. At the energy of
+about $2000$~cm$^{-1}$ above the origin the density of vibronic peaks
+increases significantly. This value can be compared to the minimum of the
+conical intersection located about $1000$~cm$^{-1}$ higher. Our
+simulation is incapable of accounting for dissociation of the molecule,
+therefore we expect a discrepancy with the experiment as the energy gets
+closer to the dissociation threshold located at $4840$~cm$^{-1}$ above the
+origin.~\cite{Willitsch:O3ZEKE:2005}
+
+Moving on to the comparison with the high-resolution spectrum, see
+Figure~\ref{fig:sim_vs_zeke}, reveals more details. Lines of the PFI-ZEKE
+experiment and our simulation show a good match. Especially the states with
+an oscillator strength originating from the A$_1$ state are well aligned with
+the experimental features. States close to the origin are similar to the
+non-coupled states. The origin peak is of a clear A$_1(0,0,0)$ character, the
+first two excitations in the symmetric bend, A$_1(0,1,0)$ and A$_1(0,2,0)$ are
+also very similar to the non-coupled states, while the higher excitations in
+this progression are showing large mixing. The same progression with one
+vibrational quantum in the symmetric stretch is more interesting. The first of
+its peaks is very well aligned with an experimentally visible feature, which
+was previously assigned as the origin of the second state. The second peak in
+this series, the first combination state A$_1(1,1,0)$, has very high
+similarity to its uncoupled version. It is a good candidate to assign the
+experimental, unassigned feature above $102,600$~cm$^{-1}$.
+
+States with an oscillator strength from the B$_2$ state are, on the other
+hand, not aligning well with the experiment. First such peak, close to the
+$12.64$~eV mark on Figure~\ref{fig:sim_vs_zeke}, is of a vibronic
+character and we assign it as a A$_1$(0,0,1) state which gains intensity
+thanks to the coupling to the B$_2$ state. This vibronic peak lies in the part
+of the spectrum marked as pollution due to water. The origin of the B$_2$
+state is very similar to the uncoupled B$_2(0,0,0)$ state, but it is located
+in an empty area of the experimental spectrum. The next peak, slightly above
+the $12.76$~eV mark on Figure~\ref{fig:sim_vs_zeke}, corresponds to the one
+excitation of the symmetric bend in the B$_2$ state, but as the label on the
+figure shows, it is only about $55$\% similar to the uncoupled state.
+
+The comparison to the earlier simulation on
+Figure~\ref{fig:sim_vs_tarronicarter} shows that both present a good match to
+the experiment. Both simulations also agree in the assignment of the first
+progression in the bending mode. The previous simulation, however, shows
+smaller spacing between the remaining lines, showing much higher congestion of
+the spectrum. Additionally the origin of the second state falls at lower
+energy aligning well with the peak that was assigned in the PFI-ZEKE
+experiment also as the origin of the second state. While an additional
+simulation of the line intensities would allow for the most complete
+comparison to the experimental spectrum, the strong alignment of the simulated
+peaks of a leading A$_1$ character with the experimental features leads us to
+believe that our simulation offers the best assignment of the spectrum to
+date.
+
 \section{Summary and Conclusions}
@@
 research presented here benefited from the NSF Center for Chemical Innovation
-Phase I (grant no. CHE-2221453) and U.S. Department of Energy, Basic Energy
+Phase I (grant no. CHE-2221453) and the U.S. Department of Energy, Basic Energy
 Sciences (grant no. DE-FG02-05ER15629).  All the authors of this research wish
@@
 
-\bibliographystyle{prf}
 \bibliography{allrefs}

### 2024-08-18 article.tex
Dr. Stanton's "part"

@@
 \begin{abstract}
-One of the most important areas of application of equation-of-motion coupled
+One of the most important areas of application for equation-of-motion coupled
 cluster (EOM-CC) theory is to the prediction, simulation and analysis of
@@
 
-Amongst the brotherhood of triatomic molecules, it cannot be argued that water
-(H$_2$O) is the most important, the most highly studied, and the most well
-understood.  Beyond H$_2$O, there are many triatomic molecules that have an
-environmental, technological or biological importance, have been subjected to
-many studied and are understood to various levels of detail.  Perhaps the most
-interesting such case is ozone (O$_3$), which has a vast number of important
-properties, a very rich history of study \cite{chappuis}, and – unlike the
-relatively simple water molecule – a profound quantum-mechanical complexity
-\cite{Babikov:anomalousOzone:2003}.  In the latter context, while we think of
-(and an NMR experiment would reveal) as two distinct kinds of oxygen Atoms in
-ozone, the full molecular Hamiltonian does not distinguish between the two
-types, with the three equivalent structures separated by a barrier that lies
-tantalizingly close to the O$_3$ $\rightarrow$ O$_2$ + O dissociation
-threshold (102.4 kJ/mol) \cite{Ruscis:ATcT:2022}. In reality, the energy
-levels of ozone all have a near triple ($e+a$) degeneracy, albeit with a
-tunneling splitting so small that it can be ignored, along with a
+Amongst the brotherhood of triatomic molecules, it cannot be denied that water
+(H$_2$O) is the most important, the most highly studied, and the most
+well-understood.  Beyond H$_2$O, there are many triatomic molecules that have
+an environmental, technological or biological importance, have been subjected
+to many studied and are understood to various levels of detail.  Perhaps the
+most interesting such case is ozone (O$_3$), which has a vast number of
+important properties, a very rich history of study \cite{chappuis}, and –
+unlike the relatively simple water molecule – a profound quantum-mechanical
+complexity \cite{Babikov:anomalousOzone:2003}.  In the latter context, while
+we think of (and an NMR experiment would reveal) as two distinct kinds of
+oxygen atoms in ozone, the full molecular Hamiltonian does not distinguish
+between them types, with the three equivalent structures separated by a
+barrier that lies tantalizingly close to the O$_3$ $\rightarrow$ O$_2$ + O
+dissociation threshold (102.4 kJ/mol) \cite{Ruscis:ATcT:2022}. In reality, the
+energy levels of ozone all have a near triple ($e+a$) degeneracy, albeit with
+a tunneling splitting so small that it can be ignored, along with a
 semi-infinite lifetime (despite opposing opinion
@@
 curious molecule.  For example, the distribution of the eighteen distinct
-$^{16}$O/$^{17}$O/$^{18}$O isotopologues in the Earth’s atmosphere differs
+$^{16}$O/$^{17}$O/$^{18}$O isotopologues in the Earth's atmosphere differs
 from what is expected based on the natural isotopic abundance, a puzzle that
@@
 
-Amongst quantum chemists, ozone has a notorious history, with its strong
+Among quantum chemists, ozone has a notorious history, with its strong
 biradical character causing significant difficulties in calculations of its
@@
 equilibrium structure of ozone would have $C_s$ symmetry (that is, the
-asymmetric stretching harmonic frequency predicted by this method was
+asymmetric stretching harmonic frequency predicted by this method is
 imaginary), a finding that led to a search for better treatment of
@@
 
-Qualitatively, the challenge posed to electronic structure by ozone ultimately
-arises from its closely spaced highest-occupied ($b_2$) and lowest-unoccupied
-($a_1$) molecular orbitals (HOMO and LUMO, respectively).  The two electron
-configurations $[\cdots]b_2^2 a_1^0$ and $[\cdots]b_2^0 a_1^2$ mix strongly,
-posing the aforementioned challenges with (especially single-reference)
-quantum chemical methods.   Another consequence of the identity and energetic
-proximity of these two orbitals is that the ozone cation (which is
-isoelectronic to the NO$_2$ radical) has closely lying $^2A_1$ and $^2B_2$
-electronic states.   Like the associated states in NO$_2$, both of these
-states are plagued by orbital symmetry-breaking problems
+Qualitatively, the challenge posed to electronic structure theory by ozone
+ultimately arises from its closely spaced highest-occupied ($a_2$) and
+lowest-unoccupied ($b_1$) molecular orbitals (HOMO and LUMO, respectively).
+The two electron configurations $[\cdots]b_2^2 a_1^0$ and $[\cdots]b_2^0
+a_1^2$ mix strongly, posing the aforementioned challenges with (especially
+single-reference) quantum chemical methods.   A second feature consequence of
+the identity and energetic proximity of these two orbitals is that the ozone
+cation (which is isoelectronic to the NO$_2$ radical) has closely lying
+$^2A_1$ and $^2B_2$ electronic states.   Like the associated states in NO$_2$,
+both of these states are plagued by orbital symmetry-breaking problems
 \cite{Davidson:SymmBreak:76}, a problem that greatly complicates
-quantum-chemical efforts to them.   One of the many accomplishments of the
-Bartlett group has been the integral role played by them in the development of
-equation-of-motion coupled-cluster theory \cite{Stanton:93:EOMCC,
-Nooijen:EOMEA:95, Bartlett:CC_review:07} (EOM-CC, also known as linear
-response coupled-cluster theory \cite{Koch:90:LinResp}). These methods provide
-a very efficient and simple way to study certain classes of what are often
-termed ``multireference problems" \cite{Krylov:EOMRev:07}, and are ideally
-suited to studying many reactive intermediates, radicals, biradicals and
-electronically excited states.   From a somewhat wider viewpoint, the
-existence of closely spaced electronic states always carries the potential for
-(possibly strong) vibronic coupling, a phenomenon that can play an important
-role in molecular dynamics and spectroscopy.  Indeed, one of the great
-successes of EOM-CC methods has been in their ability -- if and only if
-combined with vibronic coupling models -- to enable high-quality simulations
-of complicated electronic spectra.   Such work provides important insight into
-the nature of vibronic coupling in molecular systems, as has been exemplified
-by various application studies (for example, see Refs.~\cite{Stanton:NO3:07}
-and~\cite{Koppel:02}).
+quantum-chemical efforts to characterize them.   One of the many
+accomplishments of the Bartlett group has been the integral role played by
+them in the development of equation-of-motion coupled-cluster theory
+\cite{Stanton:93:EOMCC, Nooijen:EOMEA:95, Bartlett:CC_review:07} (EOM-CC, also
+known as linear response coupled-cluster theory \cite{Koch:90:LinResp}). These
+methods provide a very efficient and simple way to study certain classes of
+what are often termed ``multireference problems" \cite{Krylov:EOMRev:07}, and
+are ideally suited to studying many reactive intermediates, radicals,
+biradicals and electronically excited states.   From a somewhat wider
+viewpoint, the existence of closely spaced electronic states always carries
+the potential for (possibly strong) vibronic coupling, a phenomenon that can
+play an important role in molecular dynamics and spectroscopy.  Indeed, one of
+the great successes of EOM-CC methods has been in their ability -- if and only
+if combined with vibronic coupling models -- to enable high-quality
+simulations of complicated electronic spectra.   Such work provides important
+insight into the nature of vibronic coupling in molecular systems, as has been
+exemplified by various application studies (for example, see
+Refs.~\cite{Stanton:NO3:07} and~\cite{Koppel:02}).
 
 Our contribution to this issue paying homage to the career and accomplishments
-of R.J. Bartlett consists of an application of a vibronic coupling model
-parametrized by EOM-CC calculations to the photoelectron spectrum of ozone.
+of R.J. Bartlett consists of an application of a vibronic coupling model,
+parametrized by EOM-CC calculations, to the photoelectron spectrum of ozone.
 We trust that this combination of methodology applied to a molecule that has

### 2024-08-23 article.tex
Turn the document into single-spaces
Update to the summary from Dr. Stanton

@@
-\documentclass[12pt,a4paper,prb,aps,superscriptaddress]{revtex4}
+\documentclass[
+12pt,
+a4paper,
+prb,
+superscriptaddress,
+tightenlines,  % single-spaced
+]{revtex4}
 \usepackage{amsmath,amssymb,physics}
@@
 \usepackage{natbib} % bibliography
-\setcitestyle{super, sort&compress, comma} % chemistry style
+% \setcitestyle{super, sort&compress, comma} % chemistry style (prb's default)
 \usepackage{hyperref} % Clickable links in the document and their colors
@@
 
-We have simulated the vibronic effects in the photoelectron spectrum of the
-ozone.  We have used the KDC Hamiltonian\cite{Cederbaum:LVC:84, KDC:81,
-Koppel:CIbookCh7:04} with the \emph{ab initio} parametrisation of the
+In this exploration of the photoelectron spectrum of the ozone molecule,
+cutting-edge high-order coupled cluster models have been used together with a
+vibronic Hamiltonian approach to spectroscopy beyond the Born-Oppenheimer
+approximation \cite{Cederbaum:LVC:84, KDC:81, Koppel:CIbookCh7:04}.
+Calculations based on the equation-of-motion method known as EOMIP-CC have
+been used to parametrize the Hamiltonian, which relies importantly on the
 quasi-diabatic states of Ichino, Gauss and Stanton.~\cite{Stanton:EOMIPdeg:09}
-The results of our simulation match well the photoelectron spectrum from
-1970s~\cite{dyke:O3:74} and the PFI-ZEKE spectrum from
-2005.~\cite{Willitsch:O3ZEKE:2005} We present analysis of the spectra and
-highlight the role of the vibronic coupling effects. Our simulation offers a
-state-of-the-art insight into the spectrum and allows us to assign the
-simulated peaks. Our assignment agrees well with the previously assigned
-progression in the symmetric bend but we reassign some of the other features
-of the spectrum. Simulated peaks that gain intensity from the B$_2$ state are
-absent in the PFI-ZEKE spectrum which offers an interesting avenue for further
-investigations. Modeling of the intensity of the PFI-ZEKE peaks is not
-unlikely to bring insight into this problem.  Additionally our simulation does
-not model dissociation of the ozone cation, which will play a role in the high
-energy part of the spectrum.
+The results of the simulation provide an excellent match to the experimental
+spectrum from the ionization potential of the lower state to about 2500
+cm$^{-1}$ higher (the restriction arising from the local parametrization of
+the Hamiltonian).  Analysis of the spectrum and its simulation underscore the
+importance of vibronic coupling effects in the ozone cation.  Results of this
+work offer a state-of-the-art insight into the spectrum and allows for an
+assignment of the experimental results.  Simulated peaks that gain intensity
+from the ${\tilde A}B_2$ state are absent in the PFI-ZEKE spectrum which
+offers an interesting avenue for further investigations.
 

### 2024-09-02 article.tex
Update Resuts and Disucussion with Dr. Stanton comments

@@
 superscriptaddress,
-tightenlines,  % single-spaced
+% tightenlines,  % single-spaced
 ]{revtex4}
@@
 One of the most important areas of application for equation-of-motion coupled
-cluster (EOM-CC) theory is to the prediction, simulation and analysis of
+cluster (EOM-CC) theory is the prediction, simulation and analysis of
 various types of electronic spectra.   In this work, the EOM-CC method known
@@
 most interesting such case is ozone (O$_3$), which has a vast number of
-important properties, a very rich history of study \cite{chappuis}, and –
-unlike the relatively simple water molecule – a profound quantum-mechanical
+important properties, a very rich history of study \cite{chappuis}, and 
+unlike the relatively simple water molecule, a profound quantum-mechanical
 complexity \cite{Babikov:anomalousOzone:2003}.  In the latter context, while
@@
 We trust that this combination of methodology applied to a molecule that has
-been extensively studied by the Bartlett group is an appropriate contribution
-to this issue.
+been extensively studied by the Bartlett group is an appropriate contribution.
 
@@
     \caption{ 
-        Two lowest states of the ozone cation and normal modes of ozone.
+        Two lowest states of the ozone cation and normal modes of the ozone
+        neutral.
     }
@@
 
-Ozone is a $C_{2v}$ molecule (we are following the Mulliken's
-convention~\cite{Mulliken:55:symnot} by placing the molecule in the $yz$ plane
-and aligning the molecular symmetry axis with the $z$ direction) with three
+Ozone is a $C_{2v}$ molecule (following Mulliken's
+convention~\cite{Mulliken:55:symnot} the molecule is placed in the $yz$-plane
+and the molecular symmetry axis aligns with the $z$ direction) with three
 normal modes: symmetric stretch, symmetric bend, and asymmetric stretch.  The
-asymmetric stretch is of a $b_2$ symmetry. The two lowest electronic states of
+asymmetric stretch is of $b_2$ symmetry. The two lowest electronic states of
 the ozone cation are close in energy. The lower state is fully symmetric while
-the higher one is of the $B_2$ symmetry, the same symmetry as the asymmetric
+the higher one is of $B_2$ symmetry, the same symmetry as the asymmetric
 stretch. The close energetic separation and the matching symmetry of the
-asymmetric stretch results in significant vibronic effects appearing in the
-ozone photoelectron spectrum.  Figure~\ref{fig:ozone_intro} presents a
-graphical summary of this information.
+asymmetric stretch is accompanied by significant vibronic effects in the
+lowest two states of the ozone photoelectron.  Figure~\ref{fig:ozone_intro}
+presents a graphical summary of this information.
 
-We simulate the vibronic states of the ozone cation using the model
+The vibronic states of the ozone cation are simulated using the model
 Hamiltonian of K{\"o}ppel, Domcke, and Cederbaum (KDC
@@
 multi-state and multi-mode Hamiltonian defined in the basis of diabatic
-states.  We consider the ozone cation in the basis of two quasi-diabatic
+states. The ozone cation is considered in the basis of two quasi-diabatic
 states coupled by one mode (mode number 3). The model also includes two
@@
         E ^{(\alpha)} 
+        + \sum _{i \in \{1,2\}} 
+            \kappa ^{(\alpha)} _i Q _i 
+        + \sum _{i,j \in \{1,2\}} 
+            \kappa ^{(\alpha)} _{ij} Q _i Q _j 
+        + \sum _{i,j,k \in \{1,2\}} 
+            \kappa ^{(\alpha)} _{ijk} Q _i Q _j Q _k 
         + \sum _{i,j,k,l \in \{1,2\}} 
-            \kappa ^{(\alpha)} _i Q _i 
-            + \kappa ^{(\alpha)} _{ij} Q _i Q _j 
-            + \kappa ^{(\alpha)} _{ijk} Q _i Q _j Q _k 
-            + \kappa ^{(\alpha)} _{ijkl} Q_i Q _j Q _k Q _l.
+            \kappa ^{(\alpha)} _{ijkl} Q_i Q _j Q _k Q _l.
     \end{equation}
@@
 \end{subequations}
-The Hamiltonian parameters are expanded around the equilibrium geometry of
-ozone. $E ^{(\alpha)}$ are the vertical ionization energies calculated at that
-geometry. $Q_i$ are the dimensionless normal coordinates of ozone. $\kappa$ are
-the coefficients of expansion of the potential along the fully symmetric
-coordinates. $\lambda$ are the linear diabatic couplings.  $\omega$ are the
-harmonic frequencies of ozone.
+The vibronic Hamiltonian elements are expanded around the equilibrium geometry
+of neutral ozone. $E ^{(\alpha)}$ are the vertical ionization energies
+calculated at that geometry. $Q_i$ are the dimensionless normal coordinates of
+ozone, $\kappa$ are the coefficients of expansion of the potential along the
+fully symmetric coordinates, and $\lambda$ the linear diabatic coupling,
+$\omega$ represent the harmonic frequencies of the neutral.
 
-We find the parameters that enter the KDC Hamiltonian using \emph{ab inito}
-coupled-cluster (CC) methods and its equation-of-motion EOM-CC
+The parameters that enter the KDC Hamiltonian are computed using \emph{ab
+inito} coupled-cluster (CC) methods and its equation-of-motion EOM-CC
 extension.~\cite{Bartlett:CC_review:07, Krylov:EOMRev:07, Bartlett:Book:09,
-Christiansen:EOMRev:11, Bartlet:EOMRev:12, Krylov:OSRev} We use the CC
+Christiansen:EOMRev:11, Bartlet:EOMRev:12, Krylov:OSRev} The CC expansion is
 truncated to singles and doubles (CCSD), singles, doubles and triples (CCSDT)
 as well as singles, doubles, triples and quadruples
-(CCSDTQ).\cite{Matthews:ncc:2015}  We use the EOM-CC for ionization potential
-(EOM-IP).~\cite{StantonGauss:EOMIP:99} We use the Ichino, Gauss and Stanton
-definition of quasi-diabatic states based on the EOM-CC method
-(EOM-CC-QD)~\cite{Stanton:EOMIPdeg:09}. In all CC and EOM-CC calculations we
-leave all electrons correlated, in other words, we do not use the
-frozen-core approximation.
+(CCSDTQ).\cite{Matthews:ncc:2015}  The states of the cation are described with
+EOM-CC for ionization potential (EOMIP).~\cite{StantonGauss:EOMIP:99} The
+definition of quasi-diabatic states of Ichino, Gauss and Stanton based on the
+EOMIP-CC method (EOMIP-CC-QD) is used here~\cite{Stanton:EOMIPdeg:09}. All CC
+and EOM-CC calculations correlate all electrons, in other words, the
+frozen-core approximation is not invoked.
 
-Using CCSDT/ANO1 we optimize the geometry of ozone and compute its harmonic
-frequencies and normal coordinates.~\cite{Almlof:ANO,Almlof:ANO:1988} At the
-same geometry we compute the linear diabatic coupling $\lambda$ using
-EOM-IP-CCSD-QD/ANO1.  We use EOM-IP-CCSDT/ANO1 on a grid to find the
-expansion coefficients $\kappa$.
+The geometry of the neutral ozone and the computation of its harmonic
+frequencies and normal coordinates uses
+CCSDT/ANO1.~\cite{Almlof:ANO,Almlof:ANO:1988} 
+The linear diabatic coupling $\lambda$ is computed at the same geometry using
+EOMIP-CCSD-QD/ANO1. The expansion coefficients $\kappa$ are computed using the
+EOMIP-CCSD/ANO1 energies calculated on a grid around the same geometry.
 
 The vertical ionization energies are calculated using a composite method.  The
-base value is the complete basis set (CBS) extrapolation of the
+starting value is the complete basis set (CBS) extrapolation of the
 EOM-IP-CCSDT/cc-pCVnZ energies with n = 5, 6.~\cite{Woon:95:CCBS} These values
@@
 basis set and the relativistic correction calculated using
-EOM-IP-CCSD/cc-pwCVTZ.~\cite{Dunning:02:p(w)CVnZ} We introduce an error
-estimates to the reported vertical energies. For the error estimate of the
-extrapolated CBS value we use half of the absolute value of the difference
-between the best \emph{ab initio} value and the extrapolated value. For the
-remaining corrections we use as the error estimate half of the absolute value
-of the correction.
+EOM-IP-CCSD/cc-pwCVTZ.~\cite{Dunning:02:p(w)CVnZ} The error estimate of the
+extrapolated CBS reference values is reported using half of the absolute value
+of the difference between the best \emph{ab initio} value and the extrapolated
+value.  For the remaining corrections (quadruples correction, $\Delta Q$, and
+the relativistic correction) the error estimate is reported as half of the
+absolute value of the correction.
 
-To compare the simulated photoelectron spectrum to the experimental one, we
-use the oscillator strengths ratio A$_1$:B$_2$ of $1$:$1.35$.\cite{KDC:O3:92}
+To compare the simulated photoelectron spectrum to experiment, the
+cross-section ratio A$_1$:B$_2$ of $1$:$1.35$ is used.\cite{KDC:O3:92}
 Additionally, the stick spectrum is broadened with the Lorenzian envelopes
-normalized to the peaks' intensities
+normalized to the feature intensities
 \begin{equation}
@@
 
-We use the \textsc{xsim} program of Dr. Stanton to simulate the spectrum using
-the basis of 50 harmonic states in each mode and 6000 iterations of the
-Lanczos procedure. All remaining \emph{ab initio} calculations are completed
-using \textsc{CFOUR}.~\cite{cfour, cfour:2020}
+The spectrum is simulated using the \textsc{xsim} program in \textsc{CFOUR},
+with the basis of 50 harmonic states in each mode and 6000 iterations of the
+Lanczos procedure.~\cite{Sharma:xsim_socjt:2024} All calculations are
+completed using \textsc{CFOUR}.~\cite{cfour, cfour:2020}
 
@@
 
-The optimized geometry of ozone gives the bond lengths of $1.270$~\AA{} and
-the bond angle of $116.9^\circ$. The two symmetric normal modes have
-frequencies $1169$~cm$^{-1}$ and $724$~cm$^{-1}$, while the asymmetric stretch
-has frequency equal to $1129$~cm$^{-1}$. The computed value of the linear
-diabatic coupling constant $\lambda$ is $1394$~cm$^{-1}$. The vertical
-ionization energy for the first excited state, $E^{(1)}$, computed using the
-composite method described above is equal to $12.827$~eV. Our error estimate
-for this value is $30$~meV, see Table~\ref{tab:vertical_ionization_energy} for
-details.  We note that the convergence of the vertical ionization gap between
-the two states is much faster and we estimate this value as equal to
-$123\pm8$~meV, see Table~\ref{tab:vertical_gap} for details.
+The optimized geometry of ozone gives a bond length of $1.270$~\AA{} and a
+bond angle of $116.9^\circ$. The two symmetric normal modes have frequencies
+$1169$~cm$^{-1}$ and $724$~cm$^{-1}$, while the asymmetric stretch has a
+frequency equal to $1129$~cm$^{-1}$. The computed value of the linear diabatic
+coupling constant $\lambda$ is $1394$~cm$^{-1}$. The vertical ionization
+energy for the first excited state, $E^{(1)}$, computed using the composite
+method described above is  $12.827$~eV. Our error estimate for this value is
+$30$~meV, see Table~\ref{tab:vertical_ionization_energy} for details.  We note
+that the convergence of the vertical gap between the two states is much faster
+and this value is estimated as $123\pm8$~meV, see Table~\ref{tab:vertical_gap}
+for details.
 
@@
             \hline
-            Contribution             &  Gap    & Error estimate \\ \hline
+            Contribution             &  Vertical gap    & Error estimate \\ \hline
             EOM-CCSDT/CBS            &  108.8           & 0.9 \\
@@
         cm$^{-1}$. Stick spectrum shows positions and intensities of all
-        simulated states. Blue color corresponds to the oscillator strength
-        originating from the $^2$A$_1$ basis state while the red one
-        corresponds to $^2$B$_2$. Gray vertical lines with captions on top
+        simulated states. Colors blue and orange mark states of A$_1$ and
+        B$_2$ symmetry, respectively. Gray vertical lines with captions on top
         indicate positions of features as measured by the PFI-ZEKE
@@
         shifted to match the PFI-ZEKE experimental origin. $D_0$ marks the
-        dissociation threshold of O$_3^+$. Gray dotted line marks the energy
-        of the minimum of the conical intersection (CI).
+        dissociation threshold of O$_3^+$.~\cite{Willitsch:O3ZEKE:2005}
+        Gray dotted line marks the energy of the minimum of the conical
+        intersection (CI) located by our Hamiltonian.
     }
@@
     \caption{
-        Simulated photoelectron spectrum of ozone, but with no account given
-        to the vibronic coupling effect.
+        Simulated Franck-Condon photoelectron spectrum of ozone, vibronic
+        coupling removed.
     }
@@
 
-We would like to assign the vibronic peaks from our simulation. To this end,
-we run the simulation once again, this time, however we set the linear
-diabatic constant to zero, i.e., $\lambda = 0$. With that change we can
-reproduce the spectrum (at an equivalent level of theory) but in an artificial
-case where there is no vibronic coupling. Figure~\ref{fig:no_coupling}
-presents this spectrum. The non-coupled spectrum is easy to assign using the
-labels that mark the symmetry of the electronic state, A$_1$ or B$_2$, and the
-vibrational state ($\nu _1 \nu_2 \nu_3$), where $\nu _i$ is the number of
-quanta in mode $i$ with $i=1$ for the symmetric stretch, $i=2$ for the
-symmetric bend and $i=3$ for the asymmetric stretch. The assigned spectrum
-shows progressions in the symmetric bend. There is one such progression in
-each electronic state.  Additionally for each state, there is also another
-progression in the symmetric bend with one excitation in the symmetric
-stretch.
+The vibronic peaks are assigned using a comparison to the Franc-Condon
+simulation.  To this end, the spectrum is simulated once again, this time,
+however, setting the linear diabatic constant to zero, i.e., $\lambda = 0$.
+That change reproduces the spectrum (at an equivalent level of theory) but in
+an artificial case where there is no vibronic coupling, i.e., the
+Franck-Condon spectrum.  Figure~\ref{fig:no_coupling} presents this spectrum.
+The non-coupled spectrum is easy to assign using the labels that mark the
+symmetry of the electronic state, A$_1$ or B$_2$, and the vibrational state
+($\nu _1 \nu_2 \nu_3$), where $\nu _i$ is the number of quanta in mode $i$
+with $i=1$ for the symmetric stretch, $i=2$ for the symmetric bend and $i=3$
+for the asymmetric stretch.  The assigned spectrum shows progressions in the
+symmetric bend. There is one such progression in each electronic state.
+Additionally for each state, there is also another progression in the
+symmetric bend with one excitation in the symmetric stretch.
 
@@
 
-We decompose the vibronic states of our main simulation in the basis of the
-artificial, uncoupled states discussed in the previous paragraph.
-Figure~\ref{fig:sim_vs_zeke} shows the assigned spectrum and
-Table~\ref{tab:peak_assignment} lists the decomposition of all peaks in the
-region of up to $3000$~cm$^{-1}$ with intensities larger than $10^{-3}$. We
-compare the assigned spectrum against the high-resolution
+The vibronic states are decomposed in the basis of the artificial, uncoupled
+states discussed in the previous paragraph.  Figure~\ref{fig:sim_vs_zeke}
+shows the assigned spectrum and Table~\ref{tab:peak_assignment} lists the
+decomposition of all peaks in the region of up to $3000$~cm$^{-1}$ with
+intensities larger than $1/300$ of the maximum peak. 
+The resulting assignment is compared against the high-resolution
 pulsed-field-ionization zero-kinetic-energy (PFI-ZEKE) spectrum from
-2005.\cite{Willitsch:O3ZEKE:2005} The PFI-ZEKE spectrum is the best source of
+Ref.~\cite{Willitsch:O3ZEKE:2005}. The PFI-ZEKE spectrum is the best source of
 information on the location of the peaks, especially the origin, against which
-align our simulation. The origin of the simulated spectrum is located
-$170$~cm$^{-1}$ lower than the experimental origin which was experimentally
-observed at $101,020.5$~cm$^{-1}$.\cite{Willitsch:O3ZEKE:2005} This is within
+the origin peak of the simulation is aligned. The origin of the simulated
+spectrum is located $170$~cm$^{-1}$ lower than the experimental origin which
+was seen at $101 020.5$~cm$^{-1}$.\cite{Willitsch:O3ZEKE:2005} This is within
 our error estimate for the vertical ionization energy
@@
 
-Lastly, we compare our simulation to an earlier accurate simulation, the work
-of Tarroni and Carter from 2011.~\cite{tarroni:O3:2011}
-Figure~\ref{fig:sim_vs_tarronicarter} presents a comparison of lines from both
-simulations. The comparison includes only the lines from the earlier
-simulation that were tabulated with an assignment.
+Lastly, a comparison to an earlier accurate simulation, the work of Tarroni
+and Carter, Ref.~\cite{Willitsch:O3ZEKE:2005}, is presented on
+Figure~\ref{fig:sim_vs_tarronicarter}. The comparison includes only the lines
+from the earlier simulation that were tabulated with an assignment.
 
@@
     \label{tab:peak_assignment}
-    \begin{tabular}{|l|l|l|}
+    \begin{tabular}{|l|l|l|l|}
         \hline
-        Peak position (cm$^{-1}$) & Assignment & Eigenvector \\
+        Energy (cm$^{-1}$) & Symmetry& Assignment & Eigenvector \\
         \hline
-$     0$ &  $A_1(000)$ & $ -0.99~A_1(000)$ \\
-$   618$ &  $A_1(010)$ & $ +0.99~A_1(010)$ \\
-$   915$ &  $A_1(001)$ & \\
-$  1076$ &  $A_1(100)$ & $ -0.99~A_1(100)$ \\
-$  1222$ &  $A_1(020)$ & $ -0.97~A_1(020)$ \\
-$  1368$ &  $B_2(000)$ & $ +0.97~B_2(000)$ \\
-$  1498$ &             & $ -0.10~B_2(000)$ \\
-$  1684$ &  $A_1(110)$ & $ -0.96~A_1(110)$ \\
-$  1796$ &  $A_1(030)$ & $ -0.87~A_1(030) -0.16~A_1(110) -0.13~A_1(020) +0.13~A_1(040)$ \\
-$  1848$ &             & $ +0.31~A_1(030)$ \\
-$  1921$ &  $B_2(010)$ & $ -0.74~B_2(010) +0.18~B_2(020) +0.11~B_2(000)$ \\
-$  1977$ &             & $ +0.24~B_2(010)$ \\
-$  2085$ &             & $ -0.52~B_2(010)$ \\
-$  2132$ &             & $ +0.25~A_1(030) +0.24~A_1(040) +0.13~A_1(120) -0.13~A_1(050)$ \\
-$  2275$ &  $A_1(120)$ & $ +0.86~A_1(120) +0.12~A_1(040)$ \\
-$  2363$ &             & $ -0.62~A_1(040) +0.40~A_1(120) -0.17~A_1(030) +0.11~A_1(050)$ \\
-$  2439$ &             & $ +0.47~A_1(040) +0.12~A_1(120)$ \\
-$  2454$ &             & $ +0.53~B_2(020) +0.24~B_2(010) -0.17~B_2(030)$ \\
-$  2576$ &             & $ +0.27~B_2(020)$ \\
-$  2736$ &             & $ +0.60~B_2(020)$ \\
-$  2886$ &  $A_1(130)$ & $ -0.81~A_1(130) +0.12~A_1(050) -0.10~A_1(120)$ \\
-$  2974$ &             & $ -0.24~B_2(110) -0.16~B_2(020)$ \\
-$  3047$ &  $A_1(050)$ & $ +0.76~A_1(050) +0.13~A_1(130)$ \\
+$     0$ & $A_1$ &  $A_1(000)$ & $ -0.99~A_1(000)$ \\
+$   618$ & $A_1$ &  $A_1(010)$ & $ +0.99~A_1(010)$ \\
+$   915$ & $B_2$ &  $A_1(001)$ & \\
+$  1076$ & $A_1$ &  $A_1(100)$ & $ -0.99~A_1(100)$ \\
+$  1222$ & $A_1$ &  $A_1(020)$ & $ -0.97~A_1(020)$ \\
+$  1368$ & $B_2$ &  $B_2(000)$ & $ +0.97~B_2(000)$ \\
+$  1498$ & $B_2$ &             & $ -0.10~B_2(000)$ \\
+$  1684$ & $A_1$ &  $A_1(110)$ & $ -0.96~A_1(110)$ \\
+$  1796$ & $A_1$ &  $A_1(030)$ & $ -0.87~A_1(030) -0.16~A_1(110) -0.13~A_1(020) +0.13~A_1(040)$ \\
+$  1848$ & $A_1$ &             & $ +0.31~A_1(030)$ \\
+$  1921$ & $B_2$ &  $B_2(010)$ & $ -0.74~B_2(010) +0.18~B_2(020) +0.11~B_2(000)$ \\
+$  1977$ & $B_2$ &             & $ +0.24~B_2(010)$ \\
+$  2085$ & $B_2$ &             & $ -0.52~B_2(010)$ \\
+$  2132$ & $A_1$ &             & $ +0.25~A_1(030) +0.24~A_1(040) +0.13~A_1(120) -0.13~A_1(050)$ \\
+$  2143$ & $A_1$ &             & \\ 
+$  2275$ & $A_1$ &  $A_1(120)$ & $ +0.86~A_1(120) +0.12~A_1(040)$ \\
+$  2363$ & $A_1$ &             & $ -0.62~A_1(040) +0.40~A_1(120) -0.17~A_1(030) +0.11~A_1(050)$ \\
+$  2436$ & $A_1$ &             & $ +0.47~A_1(040) +0.12~A_1(120)$ \\
+$  2453$ & $B_2$ &  $B_2(020)$ & $ +0.53~B_2(020) +0.24~B_2(010) -0.17~B_2(030)$ \\
+$  2576$ & $B_2$ &             & $ +0.27~B_2(020)$ \\
+$  2735$ & $B_2$ &             & $ +0.60~B_2(020)$ \\
+$  2743$ & $A_1$ &             & \\ 
+$  2862$ & $A_1$ &             & \\ 
+$  2903$ & $A_1$ &  $A_1(130)$ & $ -0.81~A_1(130) +0.12~A_1(050) -0.10~A_1(120)$ \\
+$  2969$ & $A_1$ &             & \\ 
+$  2977$ & $B_2$ &             & $ -0.24~B_2(110) -0.16~B_2(020)$ \\
+$  3006$ & $B_2$ &             & \\ 
+$  3045$ & $A_1$ &  $A_1(050)$ & $ +0.76~A_1(050) +0.13~A_1(130)$ \\
         \hline
@@
 experimental spectrum presented on Figure~\ref{fig:sim_vs_dyke}. The peak A is
-known to be a hot band.~\cite{KDC:O3:92} The simulation reproduces well the
-consecutive increase in the intensities of peaks B, C, D, and E. The spacing
-between these peaks is also well reproduced. Drop in the intensity at the peak
-F is captured by the simulation. Starting from the peak G, the simulation
-shows discrepancy with the experiment. A sudden drop in the intensity past the
-peak H is not observed in the simulation. We discuss a likely source of this
-mismatch later.
+known to be a hot band, and does not appear in our ($0$~K)
+simulation.~\cite{KDC:O3:92} The simulation reproduces well the consecutive
+increase in the intensities of peaks B, C, D, and E. The spacing between these
+peaks is also well reproduced. The drop in the intensity at the peak F is
+captured by the simulation. Starting from the peak G, the simulation shows
+discrepancy with the experiment. A sudden drop in the intensity past the peak
+H is not observed in the simulation. A likely source of this mismatch is
+discussed later.
 
@@
 introduces additional insight. The spectrum shows that peaks B and C are
-almost purely of the $^2$A$_1$ character. Starting from the peak D, the
-contributions from two states are equally important. It is also clear that the
-following peaks arise as mixtures of many vibronic peaks. At the energy of
-about $2000$~cm$^{-1}$ above the origin the density of vibronic peaks
-increases significantly. This value can be compared to the minimum of the
-conical intersection located about $1000$~cm$^{-1}$ higher. Our
-simulation is incapable of accounting for dissociation of the molecule,
-therefore we expect a discrepancy with the experiment as the energy gets
-closer to the dissociation threshold located at $4840$~cm$^{-1}$ above the
-origin.~\cite{Willitsch:O3ZEKE:2005}
+almost purely of the $^2$A$_1$ electronic and $a_1$ vibrational character.
+Starting from the peak D, the contributions from two states are equally
+important. It is also clear that the following peaks arise as mixtures of many
+vibronic peaks. At the energy of about $2000$~cm$^{-1}$ above the origin the
+density of vibronic peaks increases significantly. This value can be compared
+to the minimum of the conical intersection located about $1000$~cm$^{-1}$
+higher. Our simulation is incapable of accounting for dissociation of the
+molecule, therefore we expect a discrepancy with the experiment as the energy
+gets closer to the dissociation threshold located at $4898\pm3$~cm$^{-1}$
+above the origin.~\cite{Willitsch:O3ZEKE:2005}
 
@@
 Figure~\ref{fig:sim_vs_zeke}, reveals more details. Lines of the PFI-ZEKE
-experiment and our simulation show a good match. Especially the states with
-an oscillator strength originating from the A$_1$ state are well aligned with
-the experimental features. States close to the origin are similar to the
-non-coupled states. The origin peak is of a clear A$_1(0,0,0)$ character, the
-first two excitations in the symmetric bend, A$_1(0,1,0)$ and A$_1(0,2,0)$ are
-also very similar to the non-coupled states, while the higher excitations in
-this progression are showing large mixing. The same progression with one
-vibrational quantum in the symmetric stretch is more interesting. The first of
-its peaks is very well aligned with an experimentally visible feature, which
-was previously assigned as the origin of the second state. The second peak in
-this series, the first combination state A$_1(1,1,0)$, has very high
+experiment and our simulation show a good match. Especially the states of the
+A$_1$ symmetry are well aligned with the experimental features. States close
+to the origin are similar to the non-coupled states. The origin peak is of
+A$_1(0,0,0)$ character, the first two excitations in the symmetric bend,
+A$_1(0,1,0)$ and A$_1(0,2,0)$ are also very similar to the non-coupled states,
+while the higher excitations in this progression are showing large mixing. The
+same progression with one vibrational quanta in the symmetric stretch is more
+interesting. The first of its peaks is very well aligned with experimentally
+visible feature, which was previously assigned as the origin of the second
+electronic state. The second peak in this series, A$_1(1,1,0)$, has very high
 similarity to its uncoupled version. It is a good candidate to assign the
@@
 
-States with an oscillator strength from the B$_2$ state are, on the other
-hand, not aligning well with the experiment. First such peak, close to the
-$12.64$~eV mark on Figure~\ref{fig:sim_vs_zeke}, is of a vibronic
-character and we assign it as a A$_1$(0,0,1) state which gains intensity
-thanks to the coupling to the B$_2$ state. This vibronic peak lies in the part
-of the spectrum marked as pollution due to water. The origin of the B$_2$
-state is very similar to the uncoupled B$_2(0,0,0)$ state, but it is located
-in an empty area of the experimental spectrum. The next peak, slightly above
-the $12.76$~eV mark on Figure~\ref{fig:sim_vs_zeke}, corresponds to the one
-excitation of the symmetric bend in the B$_2$ state, but as the label on the
-figure shows, it is only about $55$\% similar to the uncoupled state.
+States of the B$_2$ symmetry are, on the other hand, not aligning well with
+the experiment. The first such peak, close to the $12.64$~eV mark on
+Figure~\ref{fig:sim_vs_zeke}, is assigned to the A$_1$(0,0,1) state.  This is
+a vibronic state that gains intensity thanks to the coupling to the electronic
+$^2$B$_2$ state. This vibronic peak lies in the part of the spectrum where
+bands of oxygen (which appears as impurity in the gas sample) are visible in
+the experiment. The origin of the $^2$B$_2$ state is very similar to the
+uncoupled B$_2(0,0,0)$ state, but it is located in an empty area of the
+experimental spectrum. The next peak, slightly above a $12.76$~eV mark on
+Figure~\ref{fig:sim_vs_zeke}, corresponds to the one excitation of the
+symmetric bend in the $^2$B$_2$ state, but as the label on the figure shows,
+it is only about $55$\% similar to the uncoupled state.
 
@@
 comparison to the experimental spectrum, the strong alignment of the simulated
-peaks of a leading A$_1$ character with the experimental features leads us to
+peaks of a A$_1$ symmetry with the experimental features leads us to
 believe that our simulation offers the best assignment of the spectrum to
@@
 importance of vibronic coupling effects in the ozone cation.  Results of this
-work offer a state-of-the-art insight into the spectrum and allows for an
-assignment of the experimental results.  Simulated peaks that gain intensity
-from the ${\tilde A}B_2$ state are absent in the PFI-ZEKE spectrum which
-offers an interesting avenue for further investigations.
+work offer a state-of-the-art insight into the details of the spectrum and
+allows for an assignment of the experimental results.  Simulated peaks that
+gain intensity from the ${\tilde A}B_2$ state are absent in the PFI-ZEKE
+spectrum which offers an interesting avenue for further investigations.
 

### 2024-09-04 article.tex
Submitted: Dr. Krylov's edits

@@
-\documentclass[
-12pt,
-a4paper,
-prb,
-superscriptaddress,
+%\documentclass[12pt,a4paper,prb,superscriptaddress,
 % tightenlines,  % single-spaced
-]{revtex4}
+%]{revtex4}
+\documentclass[12pt,prb,aps]{revtex4}
 \usepackage{amsmath,amssymb,physics}
@@
 
-\author{Pawe{\l} W{\'o}jcik}
-\affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}
-\affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE
+\author{Pawe{\l} W{\'o}jcik$^a$,  Hanna Reisler$^a$, P{\'e}ter G. Szalay$^b$, Anna I. Krylov$^{a,\dagger}$, John F. Stanton$^{b,c,*}$\\
+{\small $^a$ Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}\\
+{\small $^b$  Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE
 E{\"o}tv{\"o}s Lor{\'a}nd University, P{\'a}zm{\'a}ny P{\'e}ter stny. 1/A,
-Budapest, Hungary}
-
-\author{Anna I. Krylov}
-\affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}
-
-\author{Hannah Reisler}
-\affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}
-
-\author{P{\'e}ter G. Szalay}
-\affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE
-E{\"o}tv{\"o}s Lor{\'a}nd University, P{\'a}zm{\'a}ny P{\'e}ter stny. 1/A,
-Budapest, Hungary}
-
-\author{John F. Stanton}
-\email[Electronic address: ]{jfstanton137@gmail.com}
-\affiliation{Quantum Theory Project, Departments of Chemistry and Physics, University of Florida, Gainesville, FL, USA 32611}
-\affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE
-E{\"o}tv{\"o}s Lor{\'a}nd University, P{\'a}zm{\'a}ny P{\'e}ter stny. 1/A,
-Budapest, Hungary}
+Budapest, Hungary}\\
+{\small $^c$ Quantum Theory Project, Departments of Chemistry and Physics, University of Florida, Gainesville, FL, USA 32611}\\
+{$\dagger$ krylov@usc.edu; $^*$ jfstanton137@gmail.com}}
 
@@
 One of the most important areas of application for equation-of-motion coupled
-cluster (EOM-CC) theory is the prediction, simulation and analysis of
-various types of electronic spectra.   In this work, the EOM-CC method known
-as EOMIP-CC is applied to the closely lying and coupled pair of states of the
-ozone cation -- ${\tilde X}^2A_1$ and ${\tilde A}^2B_2$ -- using sophisticated
-approaches that extend to the full singles, doubles, triples and quadruples
-model (EOMIP-CCSDTQ).   Combined with a venerable and powerful method for
-calculating vibronic spectra from the Hamiltonian produced by EOMIP-CC
-calculations, this research provides a spectrum that is in good agreement with
-the photoelectron spectrum of ozone.   An important result here is that the
+cluster (EOM-CC) theory is the prediction, simulation, and analysis of
+various types of electronic spectra. 
+In this work, the EOM-CC method for ionized states, known
+as EOM-IP-CC, is applied to the closely lying and coupled pair of states of the
+ozone cation --- ${\tilde X}^2$A$_1$ and ${\tilde A}^2$B$_2$ --- using highly accurate 
+treatments including up to the full single, double, triple, and quadruple excitations
+(EOM-IP-CCSDTQ). Combined with a venerable and powerful method for
+calculating vibronic spectra from the Hamiltonian produced by EOM-IP-CC
+calculations, the simulations yield a spectrum that is in good agreement with
+the photoelectron spectrum of ozone. Importantly,  the
 calculations suggest that the adiabatic gap separating these two electronic
@@
 simulated spectrum together with the more precise band positions of the
-experimental measurements suggests that this energy gap is 1366$\pm$65
+experimental measurements suggests that this energy gap is 1,366$\pm$65
 cm$^{-1}$.
@@
 (H$_2$O) is the most important, the most highly studied, and the most
-well-understood.  Beyond H$_2$O, there are many triatomic molecules that have
-an environmental, technological or biological importance, have been subjected
-to many studied and are understood to various levels of detail.  Perhaps the
+well-understood. Beyond H$_2$O, many triatomic molecules that have
+an environmental, technological or biological importance have been subjected
+to many studies and are understood to various levels of detail. Perhaps the
 most interesting such case is ozone (O$_3$), which has a vast number of
@@
 unlike the relatively simple water molecule, a profound quantum-mechanical
-complexity \cite{Babikov:anomalousOzone:2003}.  In the latter context, while
-we think of (and an NMR experiment would reveal) as two distinct kinds of
-oxygen atoms in ozone, the full molecular Hamiltonian does not distinguish
-between them types, with the three equivalent structures separated by a
+complexity \cite{Babikov:anomalousOzone:2003}. Structurally, while
+we think of ozone as having two distinct kinds of oxygen atoms (and an NMR experiment would reveal that), the full molecular Hamiltonian does not distinguish
+between them. Rather, it yields the three equivalent structures separated by a
 barrier that lies tantalizingly close to the O$_3$ $\rightarrow$ O$_2$ + O
-dissociation threshold (102.4 kJ/mol) \cite{Ruscis:ATcT:2022}. In reality, the
+dissociation threshold (102.4 kJ/mol) \cite{Ruscis:ATcT:2022}, as shown in 
+Fig. \ref{fig:3fold}.
+In reality, the
 energy levels of ozone all have a near triple ($e+a$) degeneracy, albeit with
@@
 
-Beyond the structural aspects of ozone, other mysteries center around this
+\begin{figure}[h!]
+\includegraphics[width = 8 cm]{./figures/OzonePES3fold.pdf}
+\caption{Two-dimensional slice of the potential energy surface of ozone in hyperspherical coordinates at the value of hyperradiuns $\rho$=4 a.u.
+(the lowest contour is at 1 eV below the dissociation limit).
+The slice shows three equivalent minima corresponding to
+C$_{2v}$ structures separated by large barriers. 
+Reproduced with permission from Ref. \citenum{Babikov:Ozone:2003}.
+    \label{fig:3fold}}
+\end{figure}
+
+Beyond the structural aspects of ozone, other mysteries surround this
 curious molecule. For example, the distribution of the eighteen distinct
@@
 has been open for more than three decades
-\cite{Mauersberger:OzoneMystery:1990}.
+\cite{Mauersberger:OzoneMystery:1990,Babikov:Ozone:2003}.
 
-Among quantum chemists, ozone has a notorious history, with its strong
-biradical character causing significant difficulties in calculations of its
-ostensibly simple ground state molecular properties.  An early 1989 study
+Among quantum chemists, ozone has a notorious history, owing to its strong
+diradical character causing significant difficulties in calculations of its
+ostensibly simple ground-state molecular properties. An early 1989 study
 \cite{Stanton:Ozone:1989, Stanton:Ozone:1989b} by the Bartlett group and
 collaborators found that the CCSD+T(CCSD) method predicted that the molecular
-equilibrium structure of ozone would have $C_s$ symmetry (that is, the
-asymmetric stretching harmonic frequency predicted by this method is
+equilibrium structure of ozone would have C$_s$ symmetry (that is, the
+asymmetric stretching harmonic frequency computed by this method is
 imaginary), a finding that led to a search for better treatment of
 non-iterative triple excitations, ultimately leading to the well-known CCSD(T)
-treatment \cite{Raghavachari:89, Urban:ccsd(t):1985, GenrefCCSD(T):93}. While
+method \cite{Raghavachari:89, Urban:ccsd(t):1985, GenrefCCSD(T):93}. While
 CCSD(T) and higher-level coupled-cluster methods available today
 \cite{kucharski:ccsdtq:1992, Kallay:CCHigh, Matthews:ncc:2015} do a good job
-in predicting the molecular vibrational potential, an elaborate multireference
-configuration interaction study by Dawes {\it et al.} has done an excellent
-job on the entire ground-state surface out to the dissociation limit
-\cite{Dawes:ozone:2013}. As such, the quantum chemical understanding and
+in describing the equilibrium structure and vibrational potential, an elaborate multireference configuration interaction calculations can describe 
+the entire ground-state surface out to the dissociation limit
+\cite{Schinke:Ozone:2001,Babikov:Ozone:2003,Dawes:ozone:2013}, facilitating calculations of spectroscopy and chemical reactions of ozone.
+As such, the quantum chemical understanding and
 fidelity for the ground electronic state of O$_3$ is now at a mature level.
+
+\begin{figure}[h!]
+\includegraphics[width = 6cm]{./figures/MOs.pdf}
+\caption{Frontier molecular orbitals of ozone; HF/6-31G*. The lowest states of the ozone 
+cation are $^2$A$_1$ and $^2$B$_2$ derived by ionization from the HOMO-1 and HOMO-2, respectively.
+    \label{fig:MOs}}
+\end{figure}
 
@@
 ultimately arises from its closely spaced highest-occupied ($a_2$) and
-lowest-unoccupied ($b_1$) molecular orbitals (HOMO and LUMO, respectively).
-The two electron configurations $[\cdots]b_2^2 a_1^0$ and $[\cdots]b_2^0
-a_1^2$ mix strongly, posing the aforementioned challenges with (especially
-single-reference) quantum chemical methods.   A second feature consequence of
-the identity and energetic proximity of these two orbitals is that the ozone
+lowest-unoccupied ($b_1$) molecular orbitals (HOMO and LUMO, respectively; see Fig. \ref{fig:MOs}), giving rise to the following wavefunction:
+\begin{eqnarray}
+X^1A_1 = C_1 [core]^6 (3a_1)^2 (2b_2)^2 (4a_1)^2  (5a_1)^2 (3b_2)^2  (1b_1)^2   (4b_2)^2 (6a_1)^2 (1a_2)^2 (2b_1)^0 + \\
+C_2
+[core]^6 (3a_1)^2 (2b_2)^2 (4a_1)^2  (5a_1)^2 (3b_2)^2  (1b_1)^2 (4b_2)^2 (6a_1)^2 (1a_2)^0
+(2b_1)^2.
+\end{eqnarray}
+%The weights of these two leading configurations are $C_1\approx$0.952 and $C_2\approx$0.005, according to the EOM-SF-CCSD (EOM-CCSD with spin-flip) calculations\cite{sfccsd}.
+%{\bf changing notations, need to check with John! This is confirmed by SF.}
+The two electron configurations $[\cdots]a_2^2 b_1^0$ and $[\cdots]a_2^0 b_1^2$ 
+%$[\cdots]b_2^2 a_1^0$ and $[\cdots]b_2^0 a_1^2$ 
+mix strongly (the weights of these two leading configurations are $C_1\approx$0.952 and $C_2\approx$0.005,
+according to the EOM-SF-CCSD (EOM-CCSD with spin-flip) calculations\cite{sfccsd}), 
+posing the aforementioned challenges with (especially
+single-reference) quantum-chemical methods.   
+A second feature consequence of
+the symmetry and energetic proximity of these two orbitals is that the ozone
 cation (which is isoelectronic to the NO$_2$ radical) has closely lying
-$^2A_1$ and $^2B_2$ electronic states.   Like the associated states in NO$_2$,
-both of these states are plagued by orbital symmetry-breaking problems
+$^2$A$_1$ and $^2$B$_2$ electronic states derived by ionization from the HOMO-1 and HOMO-2, respectively (see Fig. \ref{fig:MOs}).
+Like the associated states in NO$_2$,
+these two states are plagued by orbital symmetry breaking
 \cite{Davidson:SymmBreak:76}, a problem that greatly complicates
-quantum-chemical efforts to characterize them.   One of the many
+their quantum-chemical treatment.  One of the many
 accomplishments of the Bartlett group has been the integral role played by
@@
 known as linear response coupled-cluster theory \cite{Koch:90:LinResp}). These
-methods provide a very efficient and simple way to study certain classes of
-what are often termed ``multireference problems" \cite{Krylov:EOMRev:07}, and
-are ideally suited to studying many reactive intermediates, radicals,
-biradicals and electronically excited states.   From a somewhat wider
+methods provide a very efficient and simple way to treat certain classes of electronic structure
+that are often termed ``multireference problems" \cite{Krylov:EOMRev:07,Krylov:OSRev}, and
+are ideally suited to studying reactive intermediates, radicals,
+diradicals, and electronically excited states. From a somewhat wider
 viewpoint, the existence of closely spaced electronic states always carries
 the potential for (possibly strong) vibronic coupling, a phenomenon that can
-play an important role in molecular dynamics and spectroscopy.  Indeed, one of
-the great successes of EOM-CC methods has been in their ability -- if and only
-if combined with vibronic coupling models -- to enable high-quality
+play an important role in molecular dynamics and spectroscopy. 
+Indeed, one of
+the great successes of EOM-CC methods has been in their ability --- if and only
+if combined with vibronic coupling models --- to enable high-quality
 simulations of complicated electronic spectra.  Such work provides important
-insight into the nature of vibronic coupling in molecular systems, as has been
+insights into the nature of vibronic coupling in molecular systems, as has been
 exemplified by various application studies (for example, see
-Refs.~\cite{Stanton:NO3:07} and~\cite{Koppel:02}).
+Refs.~\citenum{Stanton:NO3:07} and~\citenum{Koppel:02}). In the case of ozone cation, 
+the two lowest electronic states separated by a small gap of $\sim$0.1 eV are coupled  
+by the asymmetric stretch of b$_2$ symmetry (Fig. \ref{fig:vib}).
+
+\begin{figure}[h!]
+\includegraphics[width = 8 cm]{./figures/vibrations.pdf}
+\caption{Three normal modes of the neutral ozone CCSDT/ANO1. Asymmetric stretching vibration of b$_2$ symmetry can couple the two lowest states of the cation ($^2$A$_1$ and $^2$B$_2$). 
+\label{fig:vib}}
+\end{figure}
+
+The photoelectron spectrum of ozone has been reported\cite{dyke:O3:74} in 1974. Later,  much higher resolution of the positions of vibronic peaks was obtained with 
+pulsed-field-ionization zero-kinetic-energy (PFI-ZEKE) spectroscopy \cite{Willitsch:O3ZEKE:2005}.
+From the theory side, the vibronic photoelectron spectrum of ozone was modeled with a vibronic Hamiltonian parameterized using multi-reference configuration interaction method
+by Tarroni and Carter\cite{Willitsch:O3ZEKE:2005}.
 
@@
 
-\begin{figure}
-    \includegraphics[width = 16 cm]{./figures/ozone_intro}
-    \caption{ 
-        Two lowest states of the ozone cation and normal modes of the ozone
-        neutral.
-    }
-    \label{fig:ozone_intro}
-\end{figure}
+%\begin{figure}
+%    \includegraphics[width = 16 cm]{./figures/ozone_intro}
+%    \caption{ 
+%        Two lowest states of the ozone cation and normal modes of the ozone neutral.
+%    }
+%    \label{fig:ozone_intro}
+%\end{figure}
 
-\section{Theory}
+\section{Theoretical models and computational details}
 
 Ozone is a $C_{2v}$ molecule (following Mulliken's
-convention~\cite{Mulliken:55:symnot} the molecule is placed in the $yz$-plane
-and the molecular symmetry axis aligns with the $z$ direction) with three
-normal modes: symmetric stretch, symmetric bend, and asymmetric stretch.  The
-asymmetric stretch is of $b_2$ symmetry. The two lowest electronic states of
-the ozone cation are close in energy. The lower state is fully symmetric while
-the higher one is of $B_2$ symmetry, the same symmetry as the asymmetric
-stretch. The close energetic separation and the matching symmetry of the
-asymmetric stretch is accompanied by significant vibronic effects in the
-lowest two states of the ozone photoelectron.  Figure~\ref{fig:ozone_intro}
-presents a graphical summary of this information.
+convention~\cite{Mulliken:55:symnot}, the molecule is placed in the $yz$-plane
+and the molecular symmetry axis aligns with the $z$-direction) with three
+normal modes: symmetric stretch, symmetric bend, and asymmetric stretch (Fig. \ref{fig:vib}).  
+The asymmetric stretch is of $b_2$ symmetry. The two lowest electronic states of
+the ozone cation, $^2$A$_1$ and $^2$B$_2$, are very close in energy---separated by the gap 
+of mere 0.123 eV (or 990 cm$^{-1}$)---and can be coupled by the b$_2$ vibration, giving rise
+to significant vibronic effects in the photoelectron spectrum.
 
-The vibronic states of the ozone cation are simulated using the model
-Hamiltonian of K{\"o}ppel, Domcke, and Cederbaum (KDC
-Hamiltonian).\cite{Cederbaum:LVC:84,KDC:81,Koppel:CIbookCh7:04} This is a
+We simulate the vibronic states of the ozone cation using the model
+Hamiltonian of K{\"o}ppel, Domcke, and Cederbaum---the KDC
+Hamiltonian\cite{Cederbaum:LVC:84,KDC:81,Koppel:CIbookCh7:04}. This is a
 multi-state and multi-mode Hamiltonian defined in the basis of diabatic
-states. The ozone cation is considered in the basis of two quasi-diabatic
-states coupled by one mode (mode number 3). The model also includes two
-symmetric modes (modes number 1 and 2)
+states. For the ozone cation, the basis comprises two quasi-diabatic
+states (related to the $^2$A$_1$ and $^2$B$_2$ states at the ground-state minimum), coupled by one mode (asymmetric 
+stretch, mode number 3). 
+The model also includes two symmetric modes (modes number 1 and 2). 
+The matrix elements of the vibronic Hamiltonian are expanded around the equilibrium geometry
+of neutral ozone using the respective dimensionless normal coordinates, 
+$Q_i$, and the corresponding harmonic frequencies, $\omega_i$.
+The overall form  of the vibronic Hamiltonian is thus: 
 \begin{subequations}
     \begin{equation}
-        H = H _0 \mathbf{1}
-        +
+        H = H _0 \mathbf{1} +
         \begin{pmatrix}
-            V ^{(1)}  & \lambda _3 Q _3\\
-            \lambda _3 Q _3 & V ^{(2)}
-        \end{pmatrix}
+            V^{(1)}  & \lambda _3 Q _3\\
+            \lambda _3 Q _3 & V^{(2)}
+        \end{pmatrix},
 \end{equation}
+where $H _0$ contains kinetic energy operator 
+and a potential energy term for the coupling mode 
     \begin{equation}
@@
         - \omega _i \frac{\partial ^2}{\partial Q _i ^2 }\right)
-        + \frac{1}{2}\omega _3 Q _3 ^2
+        + \frac{1}{2}\omega _3 Q _3 ^2,
 \end{equation}
+and  $V^{(1)}$ and $V^{(1)}$ describe the two diabatic states expanded over $Q_1$ and $Q_2$
 \begin{equation}
@@
 \end{subequations}
-The vibronic Hamiltonian elements are expanded around the equilibrium geometry
-of neutral ozone. $E ^{(\alpha)}$ are the vertical ionization energies
-calculated at that geometry. $Q_i$ are the dimensionless normal coordinates of
-ozone, $\kappa$ are the coefficients of expansion of the potential along the
-fully symmetric coordinates, and $\lambda$ the linear diabatic coupling,
-$\omega$ represent the harmonic frequencies of the neutral.
+Here $E ^{(\alpha)}$ are the vertical ionization energies
+calculated at the equilibrium geometry of the neutral, $\kappa$ are the coefficients of expansion of the potential along the fully symmetric coordinates, 
+and $\lambda$ is the linear diabatic coupling.
+Fig. \ref{fig:KDC} shows the resulting diabatic potential energy surfaces. 
+The displacements along the two symmetric normal modes are clearly visible, 
+suggesting vibrational progressions along the bend and symmetric stretch.
 
-The parameters that enter the KDC Hamiltonian are computed using \emph{ab
-inito} coupled-cluster (CC) methods and its equation-of-motion EOM-CC
-extension.~\cite{Bartlett:CC_review:07, Krylov:EOMRev:07, Bartlett:Book:09,
-Christiansen:EOMRev:11, Bartlet:EOMRev:12, Krylov:OSRev} The CC expansion is
-truncated to singles and doubles (CCSD), singles, doubles and triples (CCSDT)
-as well as singles, doubles, triples and quadruples
-(CCSDTQ).\cite{Matthews:ncc:2015}  The states of the cation are described with
-EOM-CC for ionization potential (EOMIP).~\cite{StantonGauss:EOMIP:99} The
-definition of quasi-diabatic states of Ichino, Gauss and Stanton based on the
-EOMIP-CC method (EOMIP-CC-QD) is used here~\cite{Stanton:EOMIPdeg:09}. All CC
-and EOM-CC calculations correlate all electrons, in other words, the
-frozen-core approximation is not invoked.
+\begin{figure}[h!]
+\includegraphics[width = 8 cm]{./figures/KDC_PES.png}
+\caption{Cuts of potential energy surfaces of the two diabatic states of ozone cation 
+(corresponding to the $^2$A$_1$ and $^2$B$_2$ states at the equilibrium geometry of the 
+neutral) shown as function of the two stretching modes, $Q_1$ and $Q_2$, at $Q_3$=0.
+%The displacements along the two symmetric normal modes are clearly visible. 
+% Mode labels are correct
+%Energy scale is in wavenumbers.
+\label{fig:KDC}}
+\end{figure}
 
-The geometry of the neutral ozone and the computation of its harmonic
-frequencies and normal coordinates uses
-CCSDT/ANO1.~\cite{Almlof:ANO,Almlof:ANO:1988} 
-The linear diabatic coupling $\lambda$ is computed at the same geometry using
-EOMIP-CCSD-QD/ANO1. The expansion coefficients $\kappa$ are computed using the
-EOMIP-CCSD/ANO1 energies calculated on a grid around the same geometry.
+We computed the parameters of the KDC Hamiltonian using \emph{ab
+initio} CC and EOM-CC methods.~\cite{Bartlett:CC_review:07, Krylov:EOMRev:07, Bartlett:Book:09,Christiansen:EOMRev:11, Bartlet:EOMRev:12, Krylov:OSRev} 
+We considered methods in which the CC expansion was truncated to singles and doubles (CCSD), singles, doubles and triples (CCSDT), as well as singles, doubles, triples and quadruples (CCSDTQ).\cite{Matthews:ncc:2015}
+We computed the geometry of the neutral ozone and its harmonic
+frequencies and normal coordinates with CCSDT/ANO1.~\cite{Almlof:ANO,Almlof:ANO:1988}
 
-The vertical ionization energies are calculated using a composite method.  The
+The optimized geometry of ozone has a bond length of 1.270 \AA\ and a
+bond angle of 116.9$^\circ$. The two symmetric normal modes have frequencies
+1,169~cm$^{-1}$ and 724~cm$^{-1}$, and the asymmetric stretch has a
+frequency of 1,129~cm$^{-1}$.
+
+We described the states of the cation using 
+EOM-CC for ionization energies (known as EOM-IP-CC).~\cite{StantonGauss:EOMIP:99} 
+We used the definition of quasi-diabatic states by Ichino, Gauss, and Stanton\cite{Stanton:EOMIPdeg:09}  based on the  EOM-IP-CC method (EOMIP-CC-QD). 
+We computed the linear diabatic coupling $\lambda$ and the expansion coefficients $\kappa$ at the equilibrium geometry of the neutral; $\lambda$ was computed with 
+EOM-IP-CCSD-QD/ANO1 and  $\kappa$'s were computed using the
+EOM-IP-CCSD/ANO1 energies calculated on a grid. The computed value of the linear diabatic coupling constant $\lambda$ is 1,394~cm$^{-1}$. The complete set of KDC parameters is given in the SI.
+
+We computed the vertical ionization energies using a composite method.  The
 starting value is the complete basis set (CBS) extrapolation of the
 EOM-IP-CCSDT/cc-pCVnZ energies with n = 5, 6.~\cite{Woon:95:CCBS} These values
-are augmented with two corrections: the $\Delta$Q correction in the cc-pwCVTZ
+were augmented with two corrections: the $\Delta$Q correction in the cc-pwCVTZ
 basis set and the relativistic correction calculated using
-EOM-IP-CCSD/cc-pwCVTZ.~\cite{Dunning:02:p(w)CVnZ} The error estimate of the
-extrapolated CBS reference values is reported using half of the absolute value
+EOM-IP-CCSD/cc-pwCVTZ.~\cite{Dunning:02:p(w)CVnZ} 
+The error bars were estimated as follows: 
+for the extrapolated CBS reference values, we used half of the absolute value
 of the difference between the best \emph{ab initio} value and the extrapolated
-value.  For the remaining corrections (quadruples correction, $\Delta Q$, and
-the relativistic correction) the error estimate is reported as half of the
+value, and for the remaining corrections (quadruples correction, $\Delta Q$, and
+the relativistic correction) we used half of the
 absolute value of the correction.
 
-To compare the simulated photoelectron spectrum to experiment, the
-cross-section ratio A$_1$:B$_2$ of $1$:$1.35$ is used.\cite{KDC:O3:92}
-Additionally, the stick spectrum is broadened with the Lorenzian envelopes
+To compare the simulated photoelectron spectrum to the experimental one, 
+we used the
+cross-section ratio A$_1$:B$_2$ of $1$:$1.35$.\cite{KDC:O3:92}
+Additionally, the stick spectrum was broadened with the Lorenzian envelopes
 normalized to the feature intensities
 \begin{equation}
-    f _{envelope}(x, x _i, I _i) = 
-    \frac{I_i}{(x-x _i)^2 + (\gamma/ 2) ^2}.
+    f(x, x _i, I _i) = 
+    \frac{I_i}{(x-x _i)^2 + (\gamma/ 2) ^2},
     \label{eq:lorentzian}
 \end{equation}
-$x _i$ is the position of the spectral peak, $I _i$ is its intensity and
+where $x_i$ is the position of the spectral peak, $I_i$ is its intensity and
 $\gamma$ is the peak's width.
 
-The spectrum is simulated using the \textsc{xsim} program in \textsc{CFOUR},
-with the basis of 50 harmonic states in each mode and 6000 iterations of the
-Lanczos procedure.~\cite{Sharma:xsim_socjt:2024} All calculations are
-completed using \textsc{CFOUR}.~\cite{cfour, cfour:2020}
+The spectrum was computed using the \textsc{xsim} module in \textsc{CFOUR},
+with the basis of 50 harmonic states in each mode and 6,000 iterations of the
+Lanczos procedure.~\cite{Sharma:xsim_socjt:2024} All electrons were correlated in all CC/EOM-CC calculations. All calculations were performed 
+using \textsc{CFOUR}.~\cite{cfour, cfour:2020}
 
-\section{Results}
+\section{Results and Discussion}
 
-The optimized geometry of ozone gives a bond length of $1.270$~\AA{} and a
-bond angle of $116.9^\circ$. The two symmetric normal modes have frequencies
-$1169$~cm$^{-1}$ and $724$~cm$^{-1}$, while the asymmetric stretch has a
-frequency equal to $1129$~cm$^{-1}$. The computed value of the linear diabatic
-coupling constant $\lambda$ is $1394$~cm$^{-1}$. The vertical ionization
-energy for the first excited state, $E^{(1)}$, computed using the composite
-method described above is  $12.827$~eV. Our error estimate for this value is
-$30$~meV, see Table~\ref{tab:vertical_ionization_energy} for details.  We note
+We begin by discussing the accuracy of the computed electronic states.
+The vertical ionization energy for the first excited state, $E^{(1)}$, 
+computed using the composite
+method described above, is  12.827~eV. Our error estimate for this value is
+30~meV (see Table~\ref{tab:vertical_ionization_energy} for details).  We note
 that the convergence of the vertical gap between the two states is much faster
-and this value is estimated as $123\pm8$~meV, see Table~\ref{tab:vertical_gap}
-for details.
+and this value is estimated as 123$\pm8$~meV (see Table~\ref{tab:vertical_gap}
+for details).
 
-\begin{table}
+\begin{table}[h!]
 \caption{
 Vertical ionization energies with the error estimates, eV.
-    }
-    \label{tab:vertical_ionization_energy}
+    \label{tab:vertical_ionization_energy}}
     \begin{center}  
@@
             Contribution  & $^2$A$_1$ & $^2$B$_2$ & Error estimate \\ \hline
-            CBS           & 12.872    & 12.981    & 0.02 \\
-            $\Delta$Q     & -0.037    & -0.021    & 0.02 \\
-            Relativistic  & -0.008    & -0.010    & 0.005 \\ \hline
-            Sum           & 12.827    & 12.950    & 0.03 \\ \hline
+            EOM-CCSDT/CBS & 12.872    & 12.981    & 0.02 \\
+            $\Delta$Q/pwCVTZ & -0.037    & -0.021    & 0.02 \\
+            Relativistic/CCSD/pwCVTZ  & -0.008    & -0.010    & 0.005 \\ \hline
+            Final value, eV           & 12.827    & 12.950    & 0.03 \\ \hline
         \end{tabular}
@@
 
-\begin{table}
+\begin{table}[h!]
     \caption{
-        The energy gap between vertical ionization energies of the $^2$A$_1$
-        and $^2$B$_2$ states with error estimates, meV. See caption of
-        Table~\ref{tab:vertical_ionization_energy}.
-    }
-    \label{tab:vertical_gap}
+        The vertical energy gap between the $^2$A$_1$
+        and $^2$B$_2$ states with error estimates, meV. 
+        %See caption of Table~\ref{tab:vertical_ionization_energy}.
+    \label{tab:vertical_gap}}
     \begin{center}
@@
 \begin{figure}
-\includegraphics[width = 8 cm]{figures/sim_vs_Dyke}
+\includegraphics[width = 8 cm]{figures/sim_vs_Dyke.png}
 \caption{
-    Comparison of the experimental (solid black line) and the simulated
-    spectrum. In the simulated spectra the plot is generated with peak width
-    $\gamma = 30$~meV. The simulated spectrum is shifted towards higher
-    energies by $21$~meV.
-}
-\label{fig:sim_vs_dyke}
+    Comparison of the experimental (solid black line digitized from Ref.~\citenum{dyke:O3:74}) and the simulated
+    photoelectron spectra shown as electron binding energies.
+    The simulated spectrum is blue-shifted by 21~meV and the peaks were broadened using  $\gamma$ = 30~meV.
+    Used with permission of Royal Society of Chemistry, from Ref.~\citenum{dyke:O3:74} permission conveyed through Copyright Clearance Center, Inc.
+\label{fig:sim_vs_dyke}}
 \end{figure}
@@
 Figure~\ref{fig:sim_vs_dyke} compares the simulated spectrum to the
-experimental one taken from reference~\cite{dyke:O3:74}. Our simulation allows
-for an additional element of analysis of the simulated spectrum.
-Figure~\ref{fig:ozone_overlay} presents a decomposition of the simulated
-spectrum from Figure~\ref{fig:sim_vs_dyke}. All lines that contribute to the
+lower-resolution experimental spectrum taken from Ref.~\citenum{dyke:O3:74}. 
+Peak A is
+known to be a hot band, and does not appear in our (0~K)
+simulation.~\cite{KDC:O3:92} The simulation reproduces well the consecutive
+increase in the intensities of peaks B, C, D, and E as well as the spacing between these peaks. The drop in the intensity at peak F is
+captured by the simulation. Starting from peak G, the simulation shows
+discrepancy with the experiment. A sudden drop in the intensity past peak
+H is not observed in the simulation. We discuss a likely source of this mismatch below.
+
+Our simulation allows
+for an additional element of analysis---Figure~\ref{fig:ozone_overlay} shows a decomposition of the simulated spectrum from Figure~\ref{fig:sim_vs_dyke}. 
+All lines that contribute to the
 spectrum are marked individually and are color-coded indicating which
@@
 envelope of the spectrum is also decomposed showing contributions from both
-states. Our simulation locates the minimum of the conical intersection at
-$3174$~cm$^{-1}$ above the origin ($12.92$~eV) that is marked on the figure as
-CI.
+states. Our simulation locates the minimum of the conical intersection (marked as CI) at 3,174~cm$^{-1}$ above the origin (12.92~eV).
 
-\begin{figure}
-    \includegraphics[width=10 cm]{figures/spectrum_overline}
+\begin{figure}[h!]
+%\includegraphics[width=8cm]{figures/spectrum_overline.pdf}
+\includegraphics[width=8cm]{figures/spectrum_overline.png}
 \caption{
-        Simulated photoelectron spectrum of ozone. Bottom axis shows energy
-        scale in eV. Top axis shows energy offset from the origin in
-        cm$^{-1}$. Stick spectrum shows positions and intensities of all
-        simulated states. Colors blue and orange mark states of A$_1$ and
+        Simulated photoelectron spectrum of ozone shown as electron binding energies. Bottom axis shows energy
+        scale in eV.  
+        Top axis shows energy offset from the origin in
+        cm$^{-1}$. 
+        Stick spectrum shows positions and intensities of all
+        simulated states. Blue and orange colors mark states of A$_1$ and
         B$_2$ symmetry, respectively. Gray vertical lines with captions on top
-        indicate positions of features as measured by the PFI-ZEKE
+        indicate positions of features measured by the PFI-ZEKE
         experiment.~\cite{Willitsch:O3ZEKE:2005} The simulated spectrum was
-        shifted to match the PFI-ZEKE experimental origin. $D_0$ marks the
-        dissociation threshold of O$_3^+$.~\cite{Willitsch:O3ZEKE:2005}
+        shifted to match the PFI-ZEKE experimental origin by 21~meV. 
+        $D_0$ marks the dissociation threshold of O$_3^+$.~\cite{Willitsch:O3ZEKE:2005}
         Gray dotted line marks the energy of the minimum of the conical
-        intersection (CI) located by our Hamiltonian.
-    }
-    \label{fig:ozone_overlay}
+        intersection (CI) as located by our Hamiltonian.
+    \label{fig:ozone_overlay}}
 \end{figure}
 
-\begin{figure}
-    \includegraphics[width=10cm]{figures/spectrum_assigned.pdf}
-    \caption{
-        Simulated Franck-Condon photoelectron spectrum of ozone, vibronic
+\begin{figure}[h!]
+    % \includegraphics[width=10cm]{figures/spectrum_assigned.pdf}
+    \includegraphics[width=10cm]{figures/spectrum_assigned.png}
+    \caption{Simulated Franck-Condon photoelectron spectrum of ozone with vibronic
         coupling removed.
-    }
-    \label{fig:no_coupling}
+    \label{fig:no_coupling}}
 \end{figure}
 
-The vibronic peaks are assigned using a comparison to the Franc-Condon
+We assigned the vibronic peaks using a comparison to the Franck-Condon
 simulation. To this end, the spectrum is simulated once again, this time,
-however, setting the linear diabatic constant to zero, i.e., $\lambda = 0$.
-That change reproduces the spectrum (at an equivalent level of theory) but in
-an artificial case where there is no vibronic coupling, i.e., the
-Franck-Condon spectrum.  Figure~\ref{fig:no_coupling} presents this spectrum.
+however, setting the linear diabatic constant to zero ($\lambda$= 0).
+This yields the spectra of the two electronic states 
+at an equivalent level of theory but without vibronic coupling, i.e., the
+combined Franck-Condon spectra of the two states.  
+Figure~\ref{fig:no_coupling} shows this spectrum.
 The non-coupled spectrum is easy to assign using the labels that mark the
@@
 ($\nu _1 \nu_2 \nu_3$), where $\nu _i$ is the number of quanta in mode $i$
-with $i=1$ for the symmetric stretch, $i=2$ for the symmetric bend and $i=3$
-for the asymmetric stretch.  The assigned spectrum shows progressions in the
+($i=1$ for the symmetric stretch, $i=2$ for the symmetric bend and $i=3$
+for the asymmetric stretch). The assigned spectrum shows progressions in the
 symmetric bend. There is one such progression in each electronic state.
@@
 
-\begin{figure}
-    \includegraphics[width=16 cm]{./figures/sim_vs_zeke}
+The decomposition of the spectrum presented on Figure~\ref{fig:ozone_overlay}
+provides additional insight. The spectrum shows that peaks B and C are
+almost purely of the $^2$A$_1$ electronic and $a_1$ vibrational character.
+Starting from the peak D, the contributions from two states are of equal
+magnitude. It is also clear that the following peaks comprise mixtures of many
+vibronic peaks. At the energy of about 2,000~cm$^{-1}$ above the origin the
+density of vibronic peaks increases significantly. This value can be compared
+to the minimum of the conical intersection located at about 1,000~cm$^{-1}$
+above the origin. Our Hamiltonian  is not suitable for describing dissociation of the
+molecule, therefore, we expect a discrepancy with the experiment as the energy
+gets closer to the dissociation threshold located at 4,898$\pm$3~cm$^{-1}$
+above the origin.~\cite{Willitsch:O3ZEKE:2005}
+
+\begin{figure}[h!]
+\includegraphics[width=16 cm]{./figures/sim_vs_zeke.png}
 \caption{
@@
         experiment.\cite{Willitsch:O3ZEKE:2005} The simulated spectrum was
-        shifted by $21$~meV$ = 170$~cm$^{-1}$ towards higher energies.
-    }
-    \label{fig:sim_vs_zeke}
+        blue-shifted by 21~meV = 170~cm$^{-1}$. The bottom panel is adapted from 
+        Ref. \citenum{Willitsch:O3ZEKE:2005}.
+        The bottom panel adapted with permission from Ref. \citenum{Willitsch:O3ZEKE:2005} copyright AIP Publishing LLC.
+    \label{fig:sim_vs_zeke}}
 \end{figure}
 
-The vibronic states are decomposed in the basis of the artificial, uncoupled
-states discussed in the previous paragraph.  Figure~\ref{fig:sim_vs_zeke}
+%The vibronic states are decomposed in the basis of the artificial, uncoupled states discussed in the previous paragraph.  
+Figure~\ref{fig:sim_vs_zeke}
 shows the assigned spectrum and Table~\ref{tab:peak_assignment} lists the
-decomposition of all peaks in the region of up to $3000$~cm$^{-1}$ with
-intensities larger than $1/300$ of the maximum peak. 
-The resulting assignment is compared against the high-resolution
-pulsed-field-ionization zero-kinetic-energy (PFI-ZEKE) spectrum from
-Ref.~\cite{Willitsch:O3ZEKE:2005}. The PFI-ZEKE spectrum is the best source of
+decomposition of all peaks in the region of up to 3100~cm$^{-1}$. 
+We compare the resulting assignment with the high-resolution
+PFI-ZEKE spectrum from
+Ref.~\citenum{Willitsch:O3ZEKE:2005}. The PFI-ZEKE spectrum is the best source of
 information on the location of the peaks, especially the origin, against which
 the origin peak of the simulation is aligned. The origin of the simulated
-spectrum is located $170$~cm$^{-1}$ lower than the experimental origin which
-was seen at $101 020.5$~cm$^{-1}$.\cite{Willitsch:O3ZEKE:2005} This is within
+spectrum is red-shifted by 170~cm$^{-1}$ relative to the experimental origin at 101,020.5~cm$^{-1}$.\cite{Willitsch:O3ZEKE:2005} This discrepancy is within
 our error estimate for the vertical ionization energy
-($30$~meV$=240$~cm$^{-1}$).
+(30~meV=240~cm$^{-1}$).
 
-\begin{figure}
-    \includegraphics[width=16 cm]{figures/sim_vs_TarroniCarter}
-    \caption{
-        Comparison of the simulated spectrum to an earlier accurate simulation
-        of Tarroni and Carter (assigned lines).~\cite{tarroni:O3:2011}
-    }
-    \label{fig:sim_vs_tarronicarter}
-\end{figure}
-
-Lastly, a comparison to an earlier accurate simulation, the work of Tarroni
-and Carter, Ref.~\cite{Willitsch:O3ZEKE:2005}, is presented on
-Figure~\ref{fig:sim_vs_tarronicarter}. The comparison includes only the lines
-from the earlier simulation that were tabulated with an assignment.
-
-\begin{table}
-    \renewcommand{\arraystretch}{0.8}% Tighter
-    \caption{Assignment and decomposition of the eigenvectors of the ozone
-    cation.}
-    \label{tab:peak_assignment}
-    \begin{tabular}{|l|l|l|l|}
-        \hline
-        Energy (cm$^{-1}$) & Symmetry& Assignment & Eigenvector \\
-        \hline
-$     0$ & $A_1$ &  $A_1(000)$ & $ -0.99~A_1(000)$ \\
-$   618$ & $A_1$ &  $A_1(010)$ & $ +0.99~A_1(010)$ \\
-$   915$ & $B_2$ &  $A_1(001)$ & \\
-$  1076$ & $A_1$ &  $A_1(100)$ & $ -0.99~A_1(100)$ \\
-$  1222$ & $A_1$ &  $A_1(020)$ & $ -0.97~A_1(020)$ \\
-$  1368$ & $B_2$ &  $B_2(000)$ & $ +0.97~B_2(000)$ \\
-$  1498$ & $B_2$ &             & $ -0.10~B_2(000)$ \\
-$  1684$ & $A_1$ &  $A_1(110)$ & $ -0.96~A_1(110)$ \\
-$  1796$ & $A_1$ &  $A_1(030)$ & $ -0.87~A_1(030) -0.16~A_1(110) -0.13~A_1(020) +0.13~A_1(040)$ \\
-$  1848$ & $A_1$ &             & $ +0.31~A_1(030)$ \\
-$  1921$ & $B_2$ &  $B_2(010)$ & $ -0.74~B_2(010) +0.18~B_2(020) +0.11~B_2(000)$ \\
-$  1977$ & $B_2$ &             & $ +0.24~B_2(010)$ \\
-$  2085$ & $B_2$ &             & $ -0.52~B_2(010)$ \\
-$  2132$ & $A_1$ &             & $ +0.25~A_1(030) +0.24~A_1(040) +0.13~A_1(120) -0.13~A_1(050)$ \\
-$  2143$ & $A_1$ &             & \\ 
-$  2275$ & $A_1$ &  $A_1(120)$ & $ +0.86~A_1(120) +0.12~A_1(040)$ \\
-$  2363$ & $A_1$ &             & $ -0.62~A_1(040) +0.40~A_1(120) -0.17~A_1(030) +0.11~A_1(050)$ \\
-$  2436$ & $A_1$ &             & $ +0.47~A_1(040) +0.12~A_1(120)$ \\
-$  2453$ & $B_2$ &  $B_2(020)$ & $ +0.53~B_2(020) +0.24~B_2(010) -0.17~B_2(030)$ \\
-$  2576$ & $B_2$ &             & $ +0.27~B_2(020)$ \\
-$  2735$ & $B_2$ &             & $ +0.60~B_2(020)$ \\
-$  2743$ & $A_1$ &             & \\ 
-$  2862$ & $A_1$ &             & \\ 
-$  2903$ & $A_1$ &  $A_1(130)$ & $ -0.81~A_1(130) +0.12~A_1(050) -0.10~A_1(120)$ \\
-$  2969$ & $A_1$ &             & \\ 
-$  2977$ & $B_2$ &             & $ -0.24~B_2(110) -0.16~B_2(020)$ \\
-$  3006$ & $B_2$ &             & \\ 
-$  3045$ & $A_1$ &  $A_1(050)$ & $ +0.76~A_1(050) +0.13~A_1(130)$ \\
-        \hline
-    \end{tabular}
-\end{table}
-
-\section{Discussion}
-
-We first discuss the comparison of our simulations to the lower-resolution
-experimental spectrum presented on Figure~\ref{fig:sim_vs_dyke}. The peak A is
-known to be a hot band, and does not appear in our ($0$~K)
-simulation.~\cite{KDC:O3:92} The simulation reproduces well the consecutive
-increase in the intensities of peaks B, C, D, and E. The spacing between these
-peaks is also well reproduced. The drop in the intensity at the peak F is
-captured by the simulation. Starting from the peak G, the simulation shows
-discrepancy with the experiment. A sudden drop in the intensity past the peak
-H is not observed in the simulation. A likely source of this mismatch is
-discussed later.
-
-The decomposition of the spectrum presented on Figure~\ref{fig:ozone_overlay}
-introduces additional insight. The spectrum shows that peaks B and C are
-almost purely of the $^2$A$_1$ electronic and $a_1$ vibrational character.
-Starting from the peak D, the contributions from two states are equally
-important. It is also clear that the following peaks arise as mixtures of many
-vibronic peaks. At the energy of about $2000$~cm$^{-1}$ above the origin the
-density of vibronic peaks increases significantly. This value can be compared
-to the minimum of the conical intersection located about $1000$~cm$^{-1}$
-higher. Our simulation is incapable of accounting for dissociation of the
-molecule, therefore we expect a discrepancy with the experiment as the energy
-gets closer to the dissociation threshold located at $4898\pm3$~cm$^{-1}$
-above the origin.~\cite{Willitsch:O3ZEKE:2005}
-
-Moving on to the comparison with the high-resolution spectrum, see
-Figure~\ref{fig:sim_vs_zeke}, reveals more details. Lines of the PFI-ZEKE
-experiment and our simulation show a good match. Especially the states of the
+The comparison with the high-resolution spectrum (Fig.~\ref{fig:sim_vs_zeke}) reveals more details. Lines of the PFI-ZEKE
+experiment and our simulation match very well. Especially the states of the
 A$_1$ symmetry are well aligned with the experimental features. States close
 to the origin are similar to the non-coupled states. The origin peak is of
-A$_1(0,0,0)$ character, the first two excitations in the symmetric bend,
-A$_1(0,1,0)$ and A$_1(0,2,0)$ are also very similar to the non-coupled states,
-while the higher excitations in this progression are showing large mixing. The
+A$_1$(0,0,0) character, the first two excitations in the symmetric bend,
+A$_1$(0,1,0) and A$_1$(0,2,0) are also very similar to the non-coupled states,
+whereas the higher excitations in this progression show significant mixing. The
 same progression with one vibrational quanta in the symmetric stretch is more
@@
 visible feature, which was previously assigned as the origin of the second
-electronic state. The second peak in this series, A$_1(1,1,0)$, has very high
+electronic state. The second peak in this series, A$_1$(1,1,0), has very high
 similarity to its uncoupled version. It is a good candidate to assign the
-experimental, unassigned feature above $102,600$~cm$^{-1}$.
+experimental, unassigned feature above 102,600~cm$^{-1}$.
 
-States of the B$_2$ symmetry are, on the other hand, not aligning well with
-the experiment. The first such peak, close to the $12.64$~eV mark on
+States of the B$_2$ symmetry, on the other hand, do not align well with
+the experiment. The first such peak, close to the 12.64~eV mark in
 Figure~\ref{fig:sim_vs_zeke}, is assigned to the A$_1$(0,0,1) state.  This is
-a vibronic state that gains intensity thanks to the coupling to the electronic
+a vibronic state that gains intensity owing to the coupling to the electronic
 $^2$B$_2$ state. This vibronic peak lies in the part of the spectrum where
@@
 the experiment. The origin of the $^2$B$_2$ state is very similar to the
-uncoupled B$_2(0,0,0)$ state, but it is located in an empty area of the
-experimental spectrum. The next peak, slightly above a $12.76$~eV mark on
+uncoupled B$_2$(0,0,0) state, but it is located in an empty area of the
+experimental spectrum. The next peak, slightly above a 12.76~eV mark in
 Figure~\ref{fig:sim_vs_zeke}, corresponds to the one excitation of the
 symmetric bend in the $^2$B$_2$ state, but as the label on the figure shows,
-it is only about $55$\% similar to the uncoupled state.
+it is only about 55\% similar to the uncoupled state.
 
-The comparison to the earlier simulation on
-Figure~\ref{fig:sim_vs_tarronicarter} shows that both present a good match to
+\begin{figure}
+    \includegraphics[width=16 cm]{figures/sim_vs_TarroniCarter.png}
+    \caption{Comparison of the simulated spectrum to an earlier simulation
+        by Tarroni and Carter (assigned lines).~\cite{tarroni:O3:2011}
+    \label{fig:sim_vs_tarronicarter}}
+\end{figure}
+
+Finally, Figure~\ref{fig:sim_vs_tarronicarter} compares our spectrum to
+an earlier accurate simulation using multi-reference configuration interaction 
+by Tarroni and Carter\cite{Willitsch:O3ZEKE:2005}. The comparison includes only the lines
+from the earlier simulation that were tabulated with an assignment. 
+The comparison to the earlier simulation in
+Figure~\ref{fig:sim_vs_tarronicarter} shows that both match well with
 the experiment. Both simulations also agree in the assignment of the first
 progression in the bending mode. The previous simulation, however, shows
-smaller spacing between the remaining lines, showing much higher congestion of
-the spectrum. Additionally the origin of the second state falls at lower
-energy aligning well with the peak that was assigned in the PFI-ZEKE
+smaller spacing between the remaining lines, yielding much higher congestion of
+the spectrum. Additionally, the origin of the second state falls at lower
+energy aligns well with the peak that was assigned in the PFI-ZEKE
 experiment also as the origin of the second state. While an additional
@@
 
+\begin{table}[h!]
+\renewcommand{\arraystretch}{0.8}% Tighter
+\caption{Assignment and decomposition of the eigenvectors of the ozone
+    cation. The first column shows the offset of simulated peaks from 
+    the origin in cm$^{-1}$.
+    \label{tab:peak_assignment}}
+    \begin{tabular}{|l|l|l|l|}
+        \hline
+        Offset & Symmetry& Assignment & Eigenvector \\
+        \hline
+$     0$ & $A_1$ &  $A_1(000)$ & $ +0.99~A_1(000)$ \\
+$   618$ & $A_1$ &  $A_1(010)$ & $ +0.99~A_1(010)$ \\
+$   915$ & $B_2$ &  $A_1(001)$ & \\
+$  1076$ & $A_1$ &  $A_1(100)$ & $ -0.99~A_1(100)$ \\
+$  1222$ & $A_1$ &  $A_1(020)$ & $ +0.97~A_1(020)$ \\
+$  1368$ & $B_2$ &  $B_2(000)$ & $ +0.97~B_2(000)$ \\
+$  1498$ & $B_2$ &             & $ +0.10~B_2(000)$ \\
+$  1684$ & $A_1$ &  $A_1(110)$ & $ +0.96~A_1(110)$ \\
+$  1796$ & $A_1$ &  $A_1(030)$ & $ +0.87~A_1(030) +0.16~A_1(110) +0.13~A_1(020) -0.13~A_1(040)$ \\
+$  1848$ & $A_1$ &             & $ -0.31~A_1(030)$ \\
+$  1921$ & $B_2$ &  $B_2(010)$ & $ +0.74~B_2(010) -0.18~B_2(020) -0.11~B_2(000)$ \\
+$  1977$ & $B_2$ &             & $ -0.24~B_2(010)$ \\
+$  2085$ & $B_2$ &             & $ +0.52~B_2(010)$ \\
+$  2132$ & $A_1$ &             & $ -0.25~A_1(030) -0.25~A_1(040) -0.13~A_1(120) +0.13~A_1(050)$ \\
+$  2143$ & $A_1$ &             & \\
+$  2275$ & $A_1$ &  $A_1(120)$ & $ +0.86~A_1(120) +0.12~A_1(040)$ \\
+$  2362$ & $A_1$ &             & $ -0.62~A_1(040) +0.40~A_1(120) -0.17~A_1(030) +0.11~A_1(050)$ \\
+$  2437$ & $A_1$ &             & $ -0.48~A_1(040) -0.11~A_1(120)$ \\
+$  2453$ & $B_2$ &  $B_2(020)$ & $ +0.54~B_2(020) +0.25~B_2(010) -0.17~B_2(030)$ \\
+$  2537$ & $B_2$ &             & \\
+$  2576$ & $B_2$ &             & $ +0.27~B_2(020)$ \\
+$  2671$ & $A_1$ &             & $ -0.48~A_1(040) -0.26~A_1(050) -0.19~A_1(130) -0.14~A_1(060)$ \\
+$  2735$ & $B_2$ &             & $ +0.62~B_2(020)$ \\
+$  2743$ & $A_1$ &             & \\
+$  2780$ & $B_2$ &             & $ +0.22~B_2(020)$ \\
+$  2862$ & $A_1$ &  $A_1(130)$ & $ +0.74~A_1(130) +0.13~A_1(120) -0.13~A_1(050)$ \\
+$  2903$ & $A_1$ &             & $ -0.42~A_1(130)$ \\
+$  2969$ & $A_1$ &             & $ -0.25~A_1(130) +0.17~A_1(050) +0.12~A_1(040)$ \\
+$  2978$ & $B_2$ &             & $ -0.24~B_2(110) -0.19~B_2(020) +0.12~B_2(010)$ \\
+$  3006$ & $B_2$ &             & $ -0.20~B_2(030) +0.17~B_2(110) -0.15~B_2(020)$ \\
+$  3045$ & $A_1$ &  $A_1(050)$ & $ +0.79~A_1(050)$ \\
+$  3054$ & $B_2$ &             & $ +0.29~B_2(110) -0.25~B_2(030) -0.13~B_2(020)$ \\
+        \hline
+    \end{tabular}
+\end{table}
+
 \section{Summary and Conclusions}
@@
 In this exploration of the photoelectron spectrum of the ozone molecule,
-cutting-edge high-order coupled cluster models have been used together with a
+we have deployed cutting-edge high-order CC models together with a
 vibronic Hamiltonian approach to spectroscopy beyond the Born-Oppenheimer
 approximation \cite{Cederbaum:LVC:84, KDC:81, Koppel:CIbookCh7:04}.
-Calculations based on the equation-of-motion method known as EOMIP-CC have
-been used to parametrize the Hamiltonian, which relies importantly on the
-quasi-diabatic states of Ichino, Gauss and Stanton.~\cite{Stanton:EOMIPdeg:09}
-The results of the simulation provide an excellent match to the experimental
-spectrum from the ionization potential of the lower state to about 2500
+We parameterized the Hamiltonian using EOM-IP-CC calculations  and the formalism 
+of quasi-diabatic states of Ichino, Gauss and Stanton.~\cite{Stanton:EOMIPdeg:09}
+The results of the simulation are in excellent agreement with the experimental
+spectrum from the ionization energy of the lower state to about 2,500
 cm$^{-1}$ higher (the restriction arising from the local parametrization of
-the Hamiltonian).  Analysis of the spectrum and its simulation underscore the
+the Hamiltonian).  %{\bf confused, need to discuss}
+Analysis of the spectrum and its simulation underscore the
 importance of vibronic coupling effects in the ozone cation.  Results of this
@@
 allows for an assignment of the experimental results.  Simulated peaks that
-gain intensity from the ${\tilde A}B_2$ state are absent in the PFI-ZEKE
+gain intensity from the ${\tilde A}$B$_2$ state are absent in the PFI-ZEKE
 spectrum which offers an interesting avenue for further investigations.
@@
 
-This project was initiated when three of the authors (P.W, P.G.S. and J.F.S)
+This project was initiated when three of the authors (P.W, P.G.S., and J.F.S)
 were in Budapest, where J.F.S. was serving as a recipient of the John von
-Neumann Award in STEM bestowed by the Fulbright Foundation. Additional
+Neumann Award in STEM bestowed by the Fulbright Foundation. 
+P.G.S. acknowledges support by National Research, Development and Innovation Fund (NKFIF) of Hungary (Grant No. 142634). Additional
 research presented here benefited from the NSF Center for Chemical Innovation
 Phase I (grant no. CHE-2221453) and the U.S. Department of Energy, Basic Energy
-Sciences (grant no. DE-FG02-05ER15629).  All the authors of this research wish
-Prof. Bartlett a happy ninth decade of life, and hope that he continues to
+Sciences (grant no. DE-FG02-05ER15629).  The authors wish
+Prof. Bartlett a happy ninth decade of life and hope that he continues to
 stimulate others in the field with his creative insights.
 
+\section*{Conflicts of interest}
+The authors declare the following competing financial interest(s):
+A.I.K. is the president and a part-owner of Q-Chem, Inc.
+
+\section*{Data availability}
+The data that support the findings of this study are available within the article and the associated SI.
 \clearpage
 
+\bibliographystyle{prf}
+%Will remove ozone later, already in cvs
 \bibliography{allrefs}
@@
 \end{document}
-
-% vim: tw=78 spell:

### 2024-09-04 si.tex
Submitted: Dr. Krylov's edits

\documentclass[12pt,prb,aps]{revtex4-1}
\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{verbatim}

\usepackage{xr}
\externaldocument[si-]{si}

\renewcommand\thesection{\arabic{section}}
\renewcommand\thesubsection{\thesection.\arabic{subsection}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thepage}{S\arabic{page}}

%Number the pages consecutively using the format S1, S2, etc..
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{} % sets both header and footer to nothing
\renewcommand{\headrulewidth}{0pt}
\fancyfoot[C]{\thepage}

\begin{document}

\title{Vibronic Coupling Effects in the Photoelectron Spectrum of Ozone: A
Coupled-Cluster Approach:\\
Supporting Information}

\author{Pawel Wojcik}
\affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}
%AIK: Not sure this is appropriate. If you just visited for a few weeks, does not make an affiliation.
%\affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE E{\"o}tv{\"o}s Lor{\'a}nd University, P{\'a}zm{\'a}ny P{\'e}ter stny. 1/A, Budapest, Hungary}

\author{Hanna Reisler}
\affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}

\author{P{\'e}ter G. Szalay}
\affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE
E{\"o}tv{\"o}s Lor{\'a}nd University, P{\'a}zm{\'a}ny P{\'e}ter stny. 1/A,
Budapest, Hungary}

\author{Anna I. Krylov}
\affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}

\author{John F. Stanton}
\email[Electronic address: ]{jfstanton137@gmail.com}
\affiliation{Quantum Theory Project, Departments of Chemistry and Physics, University of Florida, Gainesville, FL, USA 32611}
\affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE
E{\"o}tv{\"o}s Lor{\'a}nd University, P{\'a}zm{\'a}ny P{\'e}ter stny. 1/A,
Budapest, Hungary}

\maketitle

\renewcommand{\baselinestretch}{1.0}  %for tighter appearance
\section{xsim input}
Full set of parameters used to define the KDC Hamiltonian. The parameters are presented as an input file to the \textsc{xsim} program.
\begin{verbatim}
Units
1
Dataset
1
States
2
Modes
3
Basis Functions
50 50 50
Lanczos
6000
Transition Moment
1. 1.35 ! Muller, Koppel, Cederbaum, et al. Chem. Phys. Lett. 197, 599 (1992)
Vertical Energies, cm-1
1 103456.7 ! EOMIP-CCSDT/pCVnZ/CBS n=5,6 + dQ/pwCVTZ + relativistic/CCSD/pwCVTZ
2 104449.7 ! EOMIP-CCSDT/pCVnZ/CBS n=5,6 + dQ/pwCVTZ + relativistic/CCSD/pwCVTZ
0 0 0 0 0 0 0 0 0 0 0 0
Harmonic Frequencies, cm-1 ! CCSDT/ANO1 for the neutral state
1 724.19
2 1168.58
3 1128.50
0 0 0 0 0 0 0 0 0 0 0 0
Linear, cm-1 ! Everything below: EOM-CCSDT/ANO1
1 1 2     1178.18
1 1 1     1261.02
2 2 2     -781.75
2 2 1    -1303.82
1 2 3     1393.82
0 0 0 0 0 0 0 0 0 0 0 0
Quadratic, cm-1
1 1 2 2      802.90
1 1 1 2     -185.50
1 1 1 1      779.81
2 2 2 2     1276.61
2 2 1 2       23.56
2 2 1 1      369.33
0 0 0 0 0 0 0 0 0 0 0 0
Cubic, cm-1
1 1 2 2 2     -202.85
1 1 1 2 2       74.39
1 1 1 1 2      -62.49
1 1 1 1 1       90.19
2 2 2 2 2     -282.04
2 2 1 2 2       75.56
2 2 1 1 2       -0.49
2 2 1 1 1       71.32
0 0 0 0 0 0 0 0 0 0 0 0
Quartic, cm-1
1 1 2 2 2 2       34.17
1 1 1 2 2 2      -26.65
1 1 1 1 2 2        3.16
1 1 1 1 1 2      -12.25
1 1 1 1 1 1        3.92
2 2 2 2 2 2       41.58
2 2 1 2 2 2      -35.55
2 2 1 1 2 2       -7.14
2 2 1 1 1 2       -6.88
2 2 1 1 1 1       11.34
0 0 0 0 0 0 0 0 0 0 0 0
\end{verbatim}

\end{document}
